---
title: "Candida in Lower Respiratory Tract: Clinical Outcomes Analysis"
subtitle: "FACTS-ALIR Cohort - Publication-Ready Analysis"
author: "Georgios Kitsios, MD PhD"
date: "`r Sys.Date(7.12.26)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: cosmo
---

# Candida in Lower Respiratory Tract: Clinical Outcomes Analysis

**Publication-Ready Analysis Script**

---

## Setup

This R Markdown document performs comprehensive analysis of lower respiratory tract culture patterns and clinical outcomes in mechanically ventilated patients with acute respiratory failure.

**Dataset:** `factsalir_12.23.25.csv` (n=896 baseline observations)  
**Analysis Population:** Culture-positive patients (n=474 after filtering)  
**Analysis Date:** June 2026

---

```{r setup, include=FALSE}
################################################################################
# SECTION 0: ENVIRONMENT SETUP
################################################################################

# Set code chunk defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)

# Clear workspace
rm(list = ls())
gc()

# Set working directory
# IMPORTANT: Change this to your working directory path
working_dir <- "analysis_humans_6.2026"

# Check if directory exists, create if not
if (!dir.exists(working_dir)) {
  dir.create(working_dir, recursive = TRUE)
  cat("Created working directory:", working_dir, "\n")
}

# Set as working directory
setwd(working_dir)

cat("Working directory set to:", getwd(), "\n\n")

# Create subdirectories for organized outputs
output_dir <- "outputs"
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

if (!dir.exists(file.path(output_dir, "tables"))) {
  dir.create(file.path(output_dir, "tables"))
}

if (!dir.exists(file.path(output_dir, "figures"))) {
  dir.create(file.path(output_dir, "figures"))
}

cat("Output directories created\n")
cat("Working structure:\n")
cat("  ", working_dir, "/\n")
cat("    ├── outputs/\n")
cat("    │   ├── tables/\n")
cat("    │   └── figures/\n")
cat("    └── factsalir_12.23.25.csv\n\n")
```

---

```{r load_packages}
################################################################################
# Load Required Packages
################################################################################

# Install packages if needed (uncomment to run)
# install.packages(c("tidyverse", "survival", "survminer", "tableone", "ggpubr", "scales"))

library(tidyverse)      # Data manipulation (dplyr, ggplot2, etc.)
library(survival)       # Survival analysis functions
library(survminer)      # Visualization of survival curves
library(tableone)       # Table 1 generation (standardized approach)
library(ggpubr)         # Enhanced ggplot2 for statistical comparisons
library(scales)         # Formatting utilities
library(readxl)
cat("✓ All packages loaded successfully\n")
```



# SECTION 1: Data Loading & Initial Quality Checks

```{r load_data}
################################################################################
# Load cleaned dataset and perform basic validation
################################################################################

cat("\n")
cat("================================================================================\n")
cat("INITIAL DATA CHARACTERISTICS\n")
cat("================================================================================\n\n")

# Load main dataset
# Make sure factsalir_12.23.25.csv is in your working directory
factsalir <- read_csv("factsalir_12.23.25.csv")

# Report initial dimensions
cat(sprintf("Full factsalir dataset: %d rows × %d columns\n", 
            nrow(factsalir), ncol(factsalir)))

# Report Colonization_Groups distribution (includes missing)
cat("\nColonization_Groups distribution (full dataset):\n")
print(table(factsalir$Colonization_Groups, useNA = "ifany"))
```



# SECTION 2: Data Preparation - Filter for Culture-Positive Cases

```{r filter_data}
################################################################################
# Filter for culture-positive cases
# Include only patients with documented lower respiratory tract growth
################################################################################

cat("\n")
cat("================================================================================\n")
cat("DATA FILTERING FOR CULTURE-POSITIVE CASES\n")
cat("================================================================================\n\n")

# Step 1: Filter for culture-positive cases
factsalir_cxpos <- factsalir %>%
  filter(!Colonization_Groups %in% c(NA, "NoGrowth"))

cat(sprintf("After excluding NA and NoGrowth: n=%d patients\n", nrow(factsalir_cxpos)))

# Step 2: Check for duplicate SubjectIDs
n_unique_before <- n_distinct(factsalir_cxpos$SubjectID)
n_rows_before <- nrow(factsalir_cxpos)

factsalir_cxpos <- factsalir_cxpos %>%
  filter(!duplicated(SubjectID))

n_unique_after <- n_distinct(factsalir_cxpos$SubjectID)
n_rows_after <- nrow(factsalir_cxpos)

cat(sprintf("Duplicate SubjectID check: %d rows → %d rows (removed %d duplicates)\n",
            n_rows_before, n_rows_after, n_rows_before - n_rows_after))

# Final analysis population
cat(sprintf("\n*** FINAL ANALYSIS POPULATION: n=%d ***\n\n", n_unique_after))
```

# SECTION 3: Variable Creation - Culture Result Stratification

```{r create_variables}
################################################################################
# Create standardized culture result variable for stratification
################################################################################

cat("Culture result categorization:\n")
print(table(factsalir_cxpos$Colonization_Groups))

# Create cxres_lrt variable (culture result - lower respiratory tract)
factsalir_cxpos <- factsalir_cxpos %>%
  mutate(
    # Character version for programming
    cxres_lrt = case_when(
      Colonization_Groups == "Bacterial_Only" ~ "BacterialPathogen",
      Colonization_Groups == "NRF_Only" ~ "NRF",
      Colonization_Groups == "Yeast_Present" ~ "Candida",
      TRUE ~ NA_character_
    ),
    # Factor version for statistical tests (reference = BacterialPathogen)
    cxres_lrt_f = factor(cxres_lrt,
                        levels = c("BacterialPathogen", "NRF", "Candida"))
  )

# Validate recoding
cat("\nRecoding validation:\n")
print(table(Original = factsalir_cxpos$Colonization_Groups,
           Recoded = factsalir_cxpos$cxres_lrt,
           useNA = "ifany"))
```



# SECTION 4: Outcome Variable Creation - Liberation Success

```{r create_liberation}
################################################################################
# Define "liberation success" as successful extubation among 90-day survivors
################################################################################

cat("\n")
cat("================================================================================\n")
cat("OUTCOME VARIABLE CREATION: LIBERATION SUCCESS\n")
cat("================================================================================\n\n")

# Examine underlying variables
cat("Intubation status (raw):\n")
print(table(factsalir_cxpos$intubation_status, useNA = "ifany"))

cat("\n90-Day mortality (raw):\n")
print(table(factsalir_cxpos$Mortality90DayClean, useNA = "ifany"))

cat("\nCrosstabulation - intubation status × mortality:\n")
print(table(Intubation = factsalir_cxpos$intubation_status,
           Mortality = factsalir_cxpos$Mortality90DayClean,
           useNA = "ifany"))

# Create liberation_success variable
factsalir_cxpos <- factsalir_cxpos %>%
  mutate(
    liberation_success = case_when(
      Mortality90DayClean == 1 ~ 0,           # Death = failure to liberate
      intubation_status == 1 ~ 1,             # Extubated = success
      intubation_status == 0 ~ 0,             # Still intubated = failure
      TRUE ~ NA_real_
    ),
    # Factor version for interpretation
    liberation_success_f = factor(liberation_success,
                                 levels = c(0, 1),
                                 labels = c("Failed", "Successfully Liberated"))
  )

# Validate recoding
cat("\n--- Recoding Validation ---\n")
print(table(Mortality = factsalir_cxpos$Mortality90DayClean,
           Intubation = factsalir_cxpos$intubation_status,
           Liberation = factsalir_cxpos$liberation_success_f,
           useNA = "ifany"))

# Report outcome frequencies
cat("\nLiberation success distribution:\n")
liberation_summary <- factsalir_cxpos %>%
  group_by(cxres_lrt_f) %>%
  summarise(
    n = n(),
    n_liberated = sum(liberation_success == 1, na.rm = TRUE),
    pct_liberated = 100 * mean(liberation_success == 1, na.rm = TRUE),
    .groups = "drop"
  )

print(liberation_summary)
```

---

# SECTION 5: Data Quality Assurance

```{r data_qa}
################################################################################
# Check completeness of key outcome variables
################################################################################

cat("\n")
cat("================================================================================\n")
cat("OUTCOME VARIABLE COMPLETENESS (n=", nrow(factsalir_cxpos), ")\n")
cat("================================================================================\n\n")

completeness_check <- tibble(
  Variable = c("VFD", "timetoliberation", "liberation_success",
               "Mortality90DayClean", "RALE_mean", "spd_plasma",
               "il8", "rage", "st2", "tnfr1_final"),
  Available = c(
    sum(!is.na(factsalir_cxpos$VFD)),
    sum(!is.na(factsalir_cxpos$timetoliberation)),
    sum(!is.na(factsalir_cxpos$liberation_success)),
    sum(!is.na(factsalir_cxpos$Mortality90DayClean)),
    sum(!is.na(factsalir_cxpos$RALE_mean)),
    sum(!is.na(factsalir_cxpos$spd_plasma)),
    sum(!is.na(factsalir_cxpos$il8)),
    sum(!is.na(factsalir_cxpos$rage)),
    sum(!is.na(factsalir_cxpos$st2)),
    sum(!is.na(factsalir_cxpos$tnfr1_final))
  )
) %>%
  mutate(
    Pct_Available = sprintf("%.1f%%", 100 * Available / nrow(factsalir_cxpos)),
    Available = sprintf("%d/%d", Available, nrow(factsalir_cxpos))
  )

print(completeness_check)
```

---

# SECTION 6: PRIMARY ANALYSIS - Liberation from Mechanical Ventilation

```{r primary_analysis}
################################################################################
# Primary endpoint: Time to liberation from mechanical ventilation
# Accounts for death as competing event (treated as failure to liberate)
################################################################################

cat("\n")
cat("================================================================================\n")
cat("PRIMARY ANALYSIS: LIBERATION FROM MECHANICAL VENTILATION\n")
cat("================================================================================\n\n")

# Prepare data for survival analysis
liberation_data <- factsalir_cxpos %>%
  filter(!is.na(timetoliberation), !is.na(liberation_success), !is.na(cxres_lrt_f))

cat(sprintf("Sample size for liberation analysis: n=%d\n", nrow(liberation_data)))
cat("\nSample size by culture result:\n")
print(table(liberation_data$cxres_lrt_f))

# Fit Kaplan-Meier survival curves
liberation_fit <- survfit(Surv(timetoliberation, liberation_success) ~ cxres_lrt_f,
                          data = liberation_data)

# Print survival statistics
cat("\n--- Survival Statistics ---\n")
print(liberation_fit)

# Perform log-rank test
logrank_test <- survdiff(Surv(timetoliberation, liberation_success) ~ cxres_lrt_f,
                         data = liberation_data)

cat("\nLog-rank test:\n")
cat(sprintf("Test statistic: %.2f\n", logrank_test$chisq))
cat(sprintf("P-value: %.4f\n", 1 - pchisq(logrank_test$chisq, 1)))
```


```{r plot_liberation, fig.width=12, fig.height=8}
# Create survival curve
library(gridExtra)
library(ggpubr)

# Create survival plot - using YOUR proven approach
liberation_fit_plot <- ggsurvplot(
  liberation_fit,
  pval = TRUE,
  pval.coord = c(40, 0.3),
  fun = "event",
  censor.shape = "|",
  censor = TRUE,
  title = "Liberation from Mechanical Ventilation by Lower Respiratory Culture Result",
  legend.labs = c("Bacterial pathogen", "NRF only", "Candida present"),
  palette = c("#00A087", "#3C5488", "#E64B35"),
  risk.table = TRUE,
  tables.theme = clean_theme(),
  legend.title = "Culture Result",
  font.legend = c(11, "bold", "black"),
  font.title = c(13, "bold", "black"),
  xlab = "Days from ICU admission", 
  xlim = c(0, 90),  # ← LIMIT X-AXIS TO 90 DAYS
  break.time.by = 15,  # ← NICE SPACING
  size = 1.2)


# Display
print(liberation_fit_plot)

# Combine plot + table
gridExtra::grid.arrange(
  grobs = list(liberation_fit_plot$plot, liberation_fit_plot$table),
  heights = c(3, 1)
)

# Save exactly like your proven code
pdf(file.path(output_dir, "figures", "Figure_01_Liberation_by_CultureResult.pdf"),
    width = 11, height = 6)
print(liberation_fit_plot, newpage = FALSE)
dev.off()

cat("\n✓ Figure saved\n")

```


# SECTION 7: SECONDARY ANALYSIS - Candida Prevalence Clinical Subtype

```{r secondary_analysis}

cat("\n")
cat("================================================================================\n")
cat("CANDIDA PREVALENCE ACROSS ALL subtypes\n")
cat("================================================================================\n\n")

# Calculate Candida prevalence by disease group
candida_prev_all <- factsalir_cxpos %>%
  group_by(group) %>%
  summarise(
    Total = n(),
    Candida_n = sum(cxres_lrt == "Candida"),
    Candida_pct = 100 * Candida_n / Total,
    .groups = "drop"
  ) %>%
  arrange(desc(Candida_pct))

cat("Candida colonization prevalence by disease subtype:\n\n")
print(candida_prev_all)

# Overall Candida prevalence
overall_candida <- sum(candida_prev_all$Candida_n) / sum(candida_prev_all$Total)
cat(sprintf("\nOverall Candida prevalence: %.1f%% (%d/%d patients)\n",
           100 * overall_candida, 
           sum(candida_prev_all$Candida_n),
           sum(candida_prev_all$Total)))

# Chi-square test
cat("\n--- Chi-square Test: Disease Group × Candida Presence ---\n")
chisq_table <- table(factsalir_cxpos$group, factsalir_cxpos$cxres_lrt == "Candida")
chisq_result <- chisq.test(chisq_table)
cat(sprintf("Chi-square statistic: %.2f\n", chisq_result$statistic))
cat(sprintf("P-value: %.4f\n", chisq_result$p.value))

# Save
write.csv(candida_prev_all,
         file.path(output_dir, "Candida_Prevalence_All_Groups.csv"),
         row.names = FALSE)

cat("\n✓ Results saved\n")

```


```{r plot_candida_prevalence_all, fig.width=12, fig.height=6}

library(ggpubr)

# Sort by prevalence (highest to lowest)
candida_prev_sorted <- candida_prev_all %>%
  arrange(desc(Candida_pct)) %>%
  mutate(group = factor(group, levels = group))

# Create bar plot
fig_candida_prevalence <- candida_prev_sorted %>%
  ggplot(aes(x = reorder(group, -Candida_pct), y = Candida_pct, fill = group)) +
  geom_col(alpha = 0.8, color = "black", width = 0.7) +
  geom_text(aes(label = sprintf("%.1f%%\n(n=%d/%d)", 
                               Candida_pct, Candida_n, Total)),
            vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_y_continuous(limits = c(0, max(candida_prev_sorted$Candida_pct) * 1.15),
                     breaks = seq(0, 50, by = 10)) +
  scale_fill_viridis_d(name = "Clinical Subtype", option = "turbo") +
  labs(
    title = sprintf("Candida Prevalence by Clinical Subtype \n(Overall: %.1f%%, x² p=%.4f)", 
                   100 * overall_candida, chisq_result$p.value),
    x = "Clinical Subtype",
    y = "Candida Prevalence (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 13, hjust = 0.5),
    axis.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

print(fig_candida_prevalence)

# Save
pdf(file.path(output_dir, "figures", "Figure_02_Candida_Prevalence_All_Groups.pdf"),
    width = 12, height = 6)
print(fig_candida_prevalence)
dev.off()

cat("\n✓ Figure saved: Figure_02_Candida_Prevalence_All_Groups.pdf\n")

```

```{r plot_culture_distribution_all, fig.width=14, fig.height=7}

# Create culture result distribution across all groups
fig_culture_dist_all <- factsalir_cxpos %>%
  filter(!is.na(cxres_lrt_f), !is.na(group)) %>%
  ggplot(aes(x = reorder(group, -table(group)[group]), fill = cxres_lrt_f)) +
  geom_bar(position = "fill", alpha = 0.85, color = "black", width = 0.7) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(
    name = "Culture Result",
    values = c("BacterialPathogen" = "#00A087",
               "NRF" = "#3C5488",
               "Candida" = "#E64B35"),
    labels = c("BacterialPathogen" = "Bacterial pathogen",
               "NRF" = "NRF only",
               "Candida" = "Candida present")
  ) +
  labs(
    title = "Lower Respiratory Tract Culture Results by Clinical Subtype",
    x = "Clinical Subtype",
    y = "Proportion"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    axis.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )

print(fig_culture_dist_all)

# Save
pdf(file.path(output_dir, "figures", "Figure_03_Culture_Distribution_All_Groups.pdf"),
    width = 14, height = 7)
print(fig_culture_dist_all)
dev.off()

cat("\n✓ Figure saved: Figure_03_Culture_Distribution_All_Groups.pdf\n")

```

---

# SECTION 8: TABLE 1 - Baseline Characteristics

```{r define_variables}
################################################################################
# Define core variables for Table 1
################################################################################

vars_categorical_core <- c(
  "Gender", "Race",
  "HistDiabetes", "HistCOPD", "HistCCF", "HistCRF",
  "IsActiveNeoplasm", "HistImmunosuppression", "HistCLD", "HistPulmonaryFibrosis",
  "group", "nonpulmsepsis", "Pneumonia", "Aspiration",
  "phenotype", "SepsisType", "lunginjury",
  "day1OnABX", "steroids", "VasopressorFirstWeek",
  "Mortality30DayClean", "Mortality90DayClean"
)

vars_continuous_core <- c(
  "Age", "BMI",
  "WBC", "HGB", "Platelets", "Creatinine", "BUN", "CO2",
  "lips_score", "RALE_mean",
  "il8", "tnfr1_final", "st2", "rage", "procalcitonin",
  "il8_eta_pradj", "tnfr1_eta_pradj", "st2_eta_pradj", "rage_eta_pradj",
  "VFD", "drivingpressure_imput", "ventilatory_ratio_imput",
  "WorstPaO2FiO2Ratio_sampledate_imput"
)

cat("Checking variable availability...\n")
missing_cat <- vars_categorical_core[!vars_categorical_core %in% names(factsalir_cxpos)]
missing_con <- vars_continuous_core[!vars_continuous_core %in% names(factsalir_cxpos)]

if (length(missing_cat) > 0) {
  cat("Missing categorical variables:", paste(missing_cat, collapse = ", "), "\n")
}
if (length(missing_con) > 0) {
  cat("Missing continuous variables:", paste(missing_con, collapse = ", "), "\n")
}

if (length(missing_cat) == 0 & length(missing_con) == 0) {
  cat("✓ All variables available\n")
}
```


```{r generate_tables}

table_cat <- CreateCatTable(
  vars = vars_categorical_core,
  strata = "cxres_lrt_f",
  data = factsalir_cxpos,
  test = TRUE,
  testExact = fisher.test,
  includeNA = FALSE,
  addOverall = TRUE
)

table_con <- CreateContTable(
  vars = vars_continuous_core,
  strata = "cxres_lrt_f",
  data = factsalir_cxpos,
  addOverall = TRUE
)

cat("✓ Tables generated\n")

```

```{r print_tables}

cat("\nTABLE 1: CATEGORICAL VARIABLES\n")
print(table_cat, digits = 2, pDigits = 3, printToggle = FALSE)

cat("\nTABLE 1: CONTINUOUS VARIABLES\n")
print(table_con, digits = 1, pDigits = 3, printToggle = FALSE,
      nonnormal = vars_continuous_core)

```

```{r export_tables}

cat_output <- print(table_cat, digits = 2, pDigits = 3, printToggle = FALSE)
write.csv(cat_output, 
         file.path(output_dir, "Table1_Categorical_Characteristics.csv"))

con_output <- print(table_con, digits = 1, pDigits = 3, printToggle = FALSE,
                    nonnormal = vars_continuous_core)
write.csv(con_output,
         file.path(output_dir,  "Table1_Continuous_Characteristics.csv"))

cat("✓ Tables exported\n")

```

---

################################################################################
# OUTCOME FIGURES BY CULTURE RESULT - PAIRWISE COMPARISONS
# PNAS/Nature Publication Style - Compact, Tight Figures
# 
# Shows pairwise comparisons (Candida vs BacterialPathogen, Candida vs NRF)
# with bracket lines and p-values
################################################################################

```{r setup_outcome_figures_pairwise}

library(ggbeeswarm)  # For beeswarm points
library(ggpubr)      # For statistical annotations

# Define color palette (consistent across all figures)
culture_colors <- c(
  "BacterialPathogen" = "#00A087",
  "NRF" = "#3C5488",
  "Candida" = "#E64B35"
)

culture_labels <- c(
  "BacterialPathogen" = "Bacterial pathogen",
  "NRF" = "NRF only",
  "Candida" = "Candida present"
)

# Define pairwise comparisons (Candida vs others)
pairwise_comparisons <- list(
  c("Candida", "BacterialPathogen"),
  c("Candida", "NRF")
)

# Theme for compact figures (PNAS/Nature style)
theme_compact <- function() {
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 11, hjust = 0.5),
    axis.title = element_text(face = "bold", size = 10),
    axis.text = element_text(size = 9),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2, unit = "mm")
  )
}

cat("✓ Theme, colors, and pairwise comparisons defined\n")

```

---

## Figure 1: Ventilator-Free Days (VFD)

```{r fig_vfd_by_culture_pairwise}

fig_vfd <- factsalir_cxpos %>%
  filter(!is.na(VFD), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = VFD, fill = cxres_lrt_f)) +
  # Boxplot - narrow width
  geom_boxplot(
    width = 0.4,
    alpha = 0.7,
    outlier.shape = NA,
    color = "black",
    size = 0.5
  ) +
  # Individual points with beeswarm
  geom_beeswarm(
    aes(color = cxres_lrt_f),
    alpha = 0.4,
    size = 2,
    cex = 2,
    corral = "wrap"
  ) +
  # Pairwise comparisons with bracket lines
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  # Colors
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  # Labels and formatting
  labs(
    title = "Ventilator-Free Days by culture",
    x = "",
    y = "Ventilator-Free Days"
  ) +
  # Compact theme
  theme_compact()

print(fig_vfd)

# Save individual figure
pdf(file.path(output_dir, "figures", "Outcome_01_VFD_by_Culture.pdf"),
     width = 3, height = 4.5)
print(fig_vfd)
dev.off()

cat("✓ Figure saved: Outcome_01_VFD_by_Culture.pdf\n")

```

---

## Figure 2: Driving Pressure

```{r fig_driving_pressure_by_culture_pairwise}

fig_dp <- factsalir_cxpos %>%
  filter(!is.na(drivingpressure_imput), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = drivingpressure_imput, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, 
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2, 
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  
  labs(
    title = "Driving Pressure by culture",
    x = "",
    y = "Driving Pressure (cmH₂O)"
  ) +
  theme_compact()

print(fig_dp)

pdf(file.path(output_dir, "figures", "Outcome_02_DrivingPressure_by_Culture.pdf"),
    width = 3, height = 4.5)
print(fig_dp)
dev.off()

cat("✓ Figure saved: Outcome_02_DrivingPressure_by_Culture.pdf\n")

```

---

## Figure 3: IL-8 (Plasma)

```{r fig_il8_by_culture_pairwise}

fig_il8 <- factsalir_cxpos %>%
  filter(!is.na(il8), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = il8, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  
  labs(
    title = "IL-8 (Plasma) by Culture Result",
    x = "Lower Respiratory Culture Result",
    y = "IL-8 (pg/mL, log)"
  ) +
  theme_compact()

print(fig_il8)

pdf(file.path(output_dir, "figures", "Outcome_03_IL8_by_Culture.pdf"),
    width = 4, height = 4.5)
print(fig_il8)
dev.off()

cat("✓ Figure saved: Outcome_03_IL8_by_Culture.pdf\n")

```


```{r fig_st2_by_culture_pairwise}

fig_st2 <- factsalir_cxpos %>%
  filter(!is.na(st2), !is.na(cxres_lrt_f)) %>% filter(group!="Other") %>% 
  ggplot(aes(x = cxres_lrt_f, y = st2, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  
  labs(
    title = "Plasma sST2 by culture",
    x = "",
    y = "plasma sST2 (pg/mL, log)"
  ) +
  theme_compact()

print(fig_st2)

pdf(file.path(output_dir, "figures", "plasmaSt2_by_Culture.pdf"),
    width = 4, height = 4.5)
print(fig_st2)
dev.off()


```



## Figure SP-D (Plasma)

```{r fig_spd_by_culture_pairwise}

fig_spd <- factsalir_cxpos %>%
  filter(!is.na(spd_plasma), !is.na(cxres_lrt_f)) %>% filter(group!="Other") %>% 
  ggplot(aes(x = cxres_lrt_f, y = spd_plasma, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  
  labs(
    title = "Plasma SPD by culture",
    x = "",
    y = "plasma SPD (ng/mL, log)"
  ) +
  theme_compact()

print(fig_spd)

pdf(file.path(output_dir, "figures", "Outcome_06_SPD_by_Culture.pdf"),
     width = 3, height = 4.5)
print(fig_spd)
dev.off()

cat("✓ Figure saved: Outcome_06_SPD_by_Culture.pdf\n")

```



```{r fig_il6_tnfr1_pentraxin3_by_culture}

# IL-6
fig_il6 <- factsalir_cxpos %>%
  filter(!is.na(il6), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = il6, fill = cxres_lrt_f)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  stat_compare_means(
    comparisons = pairwise_comparisons, method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  labs(title = "IL-6 (Plasma) by Culture Result", x = "",
       y = "IL-6 (pg/mL, log)") +
  theme_compact()

print(fig_il6)
pdf(file.path(output_dir, "figures", "Outcome_IL6_by_Culture.pdf"), width = 4, height = 4.5)
print(fig_il6)
dev.off()

# TNFR1
fig_tnfr1 <- factsalir_cxpos %>%
  filter(!is.na(tnfr1_final), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = tnfr1_final, fill = cxres_lrt_f)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  stat_compare_means(
    comparisons = pairwise_comparisons, method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  labs(title = "TNFR1 (Plasma) by Culture Result", x = "",
       y = "TNFR1 (pg/mL, log)") +
  theme_compact()

print(fig_tnfr1)
pdf(file.path(output_dir, "figures", "Outcome_TNFR1_by_Culture.pdf"), width = 4, height = 4.5)
print(fig_tnfr1)
dev.off()

# Pentraxin-3
fig_ptx3 <- factsalir_cxpos %>%
  filter(!is.na(pentraxin3), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = pentraxin3, fill = cxres_lrt_f)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  stat_compare_means(
    comparisons = pairwise_comparisons, method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  labs(title = "Pentraxin-3 (ETA) by Culture Result", x = "",
       y = "Pentraxin-3 (pg/mL, log)") +
  theme_compact()

print(fig_ptx3)
pdf(file.path(output_dir, "figures", "Outcome_Pentraxin3_by_Culture.pdf"), width = 4, height = 4.5)
print(fig_ptx3)
dev.off()

cat("✓ Figures saved: IL-6, TNFR1, Pentraxin-3 by Culture\n")

```


```{r import CRP}

ALIR_EDTAPlasma_ELISA_Combined <- read_xlsx("ALIR_EDTAPlasma_ELISA_Combined.xlsx")
names(ALIR_EDTAPlasma_ELISA_Combined)
names(factsalir_cxpos)
factsalir_cxpos$SubjectID

factsalir_cxpos_crp <- merge(factsalir_cxpos,ALIR_EDTAPlasma_ELISA_Combined, by = "SubjectIDDay", all.x = TRUE )
factsalir_cxpos_crp$CRP <- factsalir_cxpos_crp$CRP_ugml


fig_crp<- factsalir_cxpos_crp %>%
  filter(!is.na(CRP), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = CRP, fill = cxres_lrt_f)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  stat_compare_means(
    comparisons = pairwise_comparisons, method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  labs(title = "plasma CRP by culture", x = "",
       y = "CRP (mcg/mL, log)") +
  theme_compact()

print(fig_crp)
pdf(file.path(output_dir, "figures", "Outcome_fig_crp_by_Culture.pdf"), width = 4, height = 4.5)
print(fig_crp)
dev.off()


```


## Figure 8: sRAGE (ETA, protein-adjusted)

```{r fig_rage_eta_by_culture_pairwise}

fig_rage_eta <- factsalir_cxpos %>%
  filter(!is.na(rage_eta_pradj), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = rage_eta_pradj, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  
  labs(
    title = "LRT sRAGE by culture",
    x = "",
    y = "sRAGE/Total Proten (pg/mL/mg), log)"
  ) +
  theme_compact()

print(fig_rage_eta)

pdf(file.path(output_dir, "figures", "Outcome_08_sRAGE_ETA_by_Culture.pdf"),
     width = 3, height = 4.5)
print(fig_rage_eta)
dev.off()

cat("✓ Figure saved: Outcome_08_sRAGE_ETA_by_Culture.pdf\n")

```

## Figure 9: sST2 (ETA, protein-adjusted)

```{r fig_st2_eta_by_culture_pairwise}

fig_st2_eta <- factsalir_cxpos %>%
  filter(!is.na(st2_eta_pradj), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = st2_eta_pradj, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  
  labs(
    title = "sST2 (ETA, Protein-adjusted)",
    x = "Lower Respiratory Culture Result",
    y = "sST2 (pg/mL, log)"
  ) +
  theme_compact()

print(fig_st2_eta)

pdf(file.path(output_dir, "figures", "Outcome_09_sST2_ETA_by_Culture.pdf"),
    width = 4, height = 4.5)
print(fig_st2_eta)
dev.off()

cat("✓ Figure saved: Outcome_09_sST2_ETA_by_Culture.pdf\n")

```

```{r fig_gdf15_eta_by_culture_pvalues}

# Calculate GDF15/protein ratio
factsalir_cxpos <- factsalir_cxpos %>%
  mutate(
    gdf15_eta_normalized = gdf15_eta_final / totProt_ETA
  )

# Create figure with consistent formatting
fig_gdf15_eta <- factsalir_cxpos %>%
  filter(!is.na(gdf15_eta_normalized), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = gdf15_eta_normalized, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  # Pairwise comparisons with NUMERIC P-VALUES
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format",
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors) +
  scale_color_manual(values = culture_colors) +
  scale_y_log10() +
  
  labs(
    title = "LRT GDF15 by culture",
    x = "",
    y = "GDF15/Total Protein (pg/mL/mg)"
  ) +
  theme_compact()

print(fig_gdf15_eta)

pdf(file.path(output_dir, "figures", "Outcome_10_GDF15_ETA_by_Culture.pdf"),
     width = 3, height = 4.5)
print(fig_gdf15_eta)
dev.off()

cat("\n✓ Figure saved: Outcome_10_GDF15_ETA_by_Culture.pdf\n")

```


```{r fig_rale_by_culture_pvalues}

# Create RALE scores figure
fig_rale <- factsalir_cxpos %>%
  filter(!is.na(RALE_mean), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = RALE_mean, fill = cxres_lrt_f)) +
  
  # Boxplot - narrow width
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  
  # Individual points with beeswarm
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2,
                cex = 2, corral = "wrap") +
  
  # Pairwise comparisons with numeric p-values
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format",
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3,
    bracket.size = 0.4
  ) +
  
  # Colors - consistent with other figures
  scale_fill_manual(values = culture_colors) +
  scale_color_manual(values = culture_colors) +
  
  # Labels and formatting
  labs(
    title = "RALE Scores by culture",
    x = "",
    y = "Mean RALE Score"
  ) +
  
  # Compact theme
  theme_compact()

print(fig_rale)

# Save figure
pdf(file.path(output_dir, "figures", "Outcome_11_RALE_by_Culture.pdf"),
    width = 3, height = 4.5)
print(fig_rale)
dev.off()

cat("\n✓ Figure saved: Outcome_11_RALE_by_Culture.pdf\n")

```

```{r fig_pao2fio2_by_culture}

fig_pao2fio2 <- factsalir_cxpos %>%
  filter(!is.na(WorstPaO2FiO2Ratio_sampledate_imput), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = WorstPaO2FiO2Ratio_sampledate_imput, fill = cxres_lrt_f)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, 
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.4, size = 2, 
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  
  labs(
    title = "PaO2/FiO2 Ratio by culture",
    x = "",
    y = "PaO2/FiO2 Ratio (lowest)"
  ) +
  theme_compact()

print(fig_pao2fio2)

ggsave(file.path(output_dir, "figures", "Outcome_PaO2FiO2_by_Culture.pdf"),
       plot = fig_pao2fio2, width = 5, height = 6, dpi = 300)

cat("✓ Figure saved: Outcome_PaO2FiO2_by_Culture.pdf\n")



```



```{r multipanel_8outcomes}

library(patchwork)

# IMPORTANT: Make sure these individual figures are already created above:
# fig_dp (Driving Pressure)
# fig_rale (RALE)
# fig_vfd (VFD)
# fig_rage_eta (sRAGE ETA)
# fig_gdf15_eta (GDF15 ETA)
# fig_spd (SP-D Plasma)

# Combine into 2×3 grid with panel labels
multipanel_outcomes <- (
  fig_dp +           # A) Driving Pressure
    fig_pao2fio2 +
  fig_rale +         # B) RALE
  fig_vfd +          # C) VFD
  fig_rage_eta +     # D) sRAGE (ETA)
  fig_gdf15_eta +    # E) GDF15 (ETA)
  fig_spd       +     # F) SP-D (Plasma)
  fig_st2
) + 
plot_layout(nrow = 2, ncol = 4, heights = c(1, 1)) +
plot_annotation(tag_levels = 'A', tag_suffix = ')',
                theme = theme(plot.tag = element_text(size = 14, face = "bold")))

# Display
print(multipanel_outcomes)

# Save as PDF (10×9 inches, publication quality)
pdf(file.path(output_dir, "figures", "Multipanel_8_Outcomes_by_Culture.pdf"),
    width = 13, height = 9)
print(multipanel_outcomes)
dev.off()

```


################################################################################
# ITS FUNGAL ABUNDANCE PLOTS - PUBLICATION READY
# 
# Purpose: Create stacked bar plots showing relative and absolute abundance
#          of lower respiratory tract fungal taxa
#
# Input: 
#   - ITS_absolute_abundance.csv (from dataframe derivation)
#   - ITS_relative_abundance.csv (from dataframe derivation)
#
# Output:
#   - Figure_ITS_RelativeAbundance_3Categories.pdf (main manuscript)
#   - Figure_ITS_AbsoluteAbundance_Top20.pdf (supplementary)
#   - Figure_ITS_AbsoluteAbundance_with_Threshold.pdf (supplementary)
#
# NOTE: This code is posted publicly on GitHub
################################################################################

```{r ITS_plots_setup}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║              ITS ABUNDANCE PLOTS - LOAD AND PREPARE DATA                  ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Load required packages
library(tidyverse)
library(scales)
library(Polychrome)

# Load exported dataframes from ITS derivation pipeline
absolute_abundance <- read_csv("ITS_absolute_abundance.csv")
relative_abundance <- read_csv("ITS_relative_abundance.csv")

cat(sprintf("✓ Loaded absolute abundance: %d samples × %d variables\n", 
            nrow(absolute_abundance), ncol(absolute_abundance)))
cat(sprintf("✓ Loaded relative abundance: %d samples × %d variables\n",
            nrow(relative_abundance), ncol(relative_abundance)))

# Verify data integrity
if (nrow(absolute_abundance) != nrow(relative_abundance)) {
  stop("ERROR: Absolute and relative abundance have different numbers of samples!")
}

if (!all(absolute_abundance$SampleID == relative_abundance$SampleID)) {
  stop("ERROR: Sample order differs between absolute and relative abundance!")
}

cat("\n✓ Data integrity check passed\n")

```

################################################################################
# FIGURE 1: RELATIVE ABUNDANCE - 3 CATEGORIES (MAIN MANUSCRIPT)
################################################################################
```{r ITS_figure1_relative_3cat_CORRECT}

relative_abundance <- read_csv("ITS_relative_abundance.csv")

# Use INDIVIDUAL TAXA percentage columns (not the _reads_pct columns)
# Calculate: Candida albicans, Other Candida, Other Taxa as percentages

relative_abundance_3cat <- relative_abundance %>%
  mutate(
    # Sum all individual Candida species percentages
    Candida_albicans_cat = Candida_albicans_pct,
    
    # Other Candida = sum of all other Candida species
    Other_Candida_cat = Candida_dubliniensis_pct + 
                        Candida_tropicalis_pct + 
                        Candida_hyderabadensis_pct +
                        Candida_orthopsilosis_pct +
                        Candida_metapsilosis_pct +
                        Meyerozyma_caribbica_pct +
                        Candida_sojae_pct,
    
    # Other Taxa = everything else
    Other_Taxa = 100 - Candida_albicans_cat - Other_Candida_cat
  ) %>%
  dplyr::select(SampleID, Total_Reads, 
         Candida_albicans_cat, Other_Candida_cat, Other_Taxa,
         Calbicans_present, Calbicans_burden)

# Verify
relative_abundance_3cat <- relative_abundance_3cat %>%
  mutate(Check_Sum = Candida_albicans_cat + Other_Candida_cat + Other_Taxa)

cat("Verification: All samples sum to 100%:", 
    all(abs(relative_abundance_3cat$Check_Sum - 100) < 0.01), "\n\n")

# Convert to long format
relative_abundance_3cat_long <- relative_abundance_3cat %>%
  dplyr::select(SampleID, Candida_albicans_cat, Other_Candida_cat, Other_Taxa) %>%
  pivot_longer(
    cols = c(Candida_albicans_cat, Other_Candida_cat, Other_Taxa),
    names_to = "Category",
    values_to = "Relative_Abundance"
  ) %>%
  mutate(
    Category = case_when(
      Category == "Candida_albicans_cat" ~ "C. albicans",
      Category == "Other_Candida_cat" ~ "Other Candida spp.",
      Category == "Other_Taxa" ~ "Other Taxa"
    ),
    Category = factor(Category, 
                     levels = c("C. albicans", "Other Candida spp.", "Other Taxa"))
  )

# Order by C. albicans
ordering <- relative_abundance_3cat %>%
  arrange(desc(Candida_albicans_cat)) %>%
  pull(SampleID)

relative_abundance_3cat_long$SampleID <- factor(relative_abundance_3cat_long$SampleID, 
                                                levels = ordering)

# Create figure
fig_relative_3cat <- ggplot(relative_abundance_3cat_long, 
                            aes(x = SampleID, y = Relative_Abundance, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  labs(x = "Individual Samples", y = "Relative Abundance (%)", fill = "") +
  scale_y_continuous(labels = percent_format(scale = 1), limits = c(0, 100)) +
  scale_fill_manual(values = c(
    "C. albicans" = "brown3",
    "Other Candida spp." = "pink2",
    "Other Taxa" = "gray70"
  )) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.title = element_text(face = "bold", size = 14)) +
  ggtitle("Candida Relative Abundance in ETA Specimens by ITS Sequencing")

print(fig_relative_3cat)

ggsave("Figure_ITS_RelativeAbundance_3Categories.pdf", fig_relative_3cat,
       width = 14, height = 6.5, dpi = 300)

```

################################################################################
# ITS RELATIVE ABUNDANCE - TOP 13 TAXA PLOT (SUPPLEMENTARY)
# 
# Purpose: Stacked bar plot showing top 13 taxa relative abundances
# Includes all sample-dominant taxa (Meyerozyma, Aspergillus, Ganoderma, etc.)
# Output: Figure_ITS_RelativeAbundance_Top13Taxa.pdf
################################################################################

```{r ITS_relative_top13_setup}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║     ITS RELATIVE ABUNDANCE - TOP 13 TAXA (SUPPLEMENTARY FIGURE)           ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Load data
#relative_abundance <- read_csv("ITS_relative_abundance.csv")

cat(sprintf("Loaded: %d samples × %d columns\n\n", nrow(relative_abundance), ncol(relative_abundance)))

```

```{r ITS_relative_top13_identify_taxa}

# Identify the individual taxa percentage columns
individual_taxa_cols <- names(relative_abundance)[
  !names(relative_abundance) %in% 
  c("SampleID", "Total_Reads", "SubjectID",
    "Candida_albicans_reads_pct", "Candida_albicans_complex_reads_pct",
    "Candida_glabrata_reads_pct", "Any_Candida_reads_pct", 
    "Other_Candida_reads_pct",
    "Calbicans_present", "Any_Candida_present", "Calbicans_burden")
]

cat("Individual taxa columns identified:", length(individual_taxa_cols), "\n\n")

```

```{r ITS_relative_top13_rank_taxa}

# Calculate total relative abundance for each taxon across all samples
taxa_totals <- relative_abundance %>%
  dplyr::select(all_of(individual_taxa_cols)) %>%
  summarise(across(everything(), ~sum(., na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "Taxon", values_to = "Total_Abundance") %>%
  arrange(desc(Total_Abundance)) %>%
  mutate(
    Rank = row_number(),
    Cumulative_Abundance = cumsum(Total_Abundance),
    Cumulative_Pct = round(100 * Cumulative_Abundance / sum(Total_Abundance), 1)
  )

cat("Top 15 taxa by total abundance:\n\n")
print(head(taxa_totals, 15))

cat(sprintf("\n\nTop 13 taxa account for %.1f%% of total abundance\n", 
            taxa_totals$Cumulative_Pct[13]))

```

```{r ITS_relative_top13_select}

# Select top 13 taxa (optimized to include all sample-dominant taxa)
n_taxa_to_use <- 13

selected_taxa <- taxa_totals$Taxon[1:n_taxa_to_use]

cat("\n\n=== TOP 13 TAXA SELECTED ===\n\n")
for (i in seq_along(selected_taxa)) {
  abund <- taxa_totals$Total_Abundance[i]
  pct <- taxa_totals$Cumulative_Pct[i]
  cat(sprintf("%2d. %-35s (%.1f%% cumulative)\n", i, selected_taxa[i], pct))
}

```

```{r ITS_relative_top13_prepare_data}

# Create dataframe with top 13 taxa + Other category
relative_abundance_top13 <- relative_abundance %>%
  dplyr::select(SampleID, Total_Reads, all_of(selected_taxa)) %>%
  mutate(
    # Sum of top 13
    Top13_Sum = rowSums(across(all_of(selected_taxa)), na.rm = TRUE),
    # Other = remainder to 100%
    Other = 100 - Top13_Sum
  )

cat("\n\nSample calculations (first 5):\n")
print(relative_abundance_top13 %>% 
  head(5) %>%
dplyr::  select(SampleID, Top13_Sum, Other))

# Verify sums to 100%
relative_abundance_top13 <- relative_abundance_top13 %>%
  mutate(Check_Sum = Top13_Sum + Other)

cat("\n\nVerification - samples sum to 100%:", all(abs(relative_abundance_top13$Check_Sum - 100) < 0.01), "\n")
cat("Range:", round(min(relative_abundance_top13$Check_Sum), 2), "-", 
    round(max(relative_abundance_top13$Check_Sum), 2), "\n\n")

```

```{r ITS_relative_top13_long_format}

# Convert to long format for plotting
relative_abundance_top13_long <- relative_abundance_top13 %>%
  dplyr::select(SampleID, all_of(selected_taxa), Other) %>%
  pivot_longer(
    cols = c(all_of(selected_taxa), Other),
    names_to = "Taxon",
    values_to = "Relative_Abundance"
  )

cat(sprintf("Long format prepared: %d rows (201 samples × 14 categories)\n\n", 
            nrow(relative_abundance_top13_long)))

```

```{r medically_relevant_fungi_abundance}

# Summary for all medically relevant fungi
medically_relevant <- relative_abundance_top13 %>%
  summarise(
    # C. albicans
    Calbicans_median = median(Candida_albicans_pct, na.rm=TRUE),
    Calbicans_IQR_low = quantile(Candida_albicans_pct, 0.25, na.rm=TRUE),
    Calbicans_IQR_high = quantile(Candida_albicans_pct, 0.75, na.rm=TRUE),
    Calbicans_detected = 100 * mean(Candida_albicans_pct > 0, na.rm=TRUE),
    
    # C. dubliniensis
    Cdub_median = median(Candida_dubliniensis_pct, na.rm=TRUE),
    Cdub_detected = 100 * mean(Candida_dubliniensis_pct > 0, na.rm=TRUE),
    
    # C. tropicalis
    Ctrop_median = median(Candida_tropicalis_pct, na.rm=TRUE),
    Ctrop_detected = 100 * mean(Candida_tropicalis_pct > 0, na.rm=TRUE),
    
    # Nakaseomyces (C. glabrata complex)
    Naka_median = median(Nakaseomyces_Genus_pct, na.rm=TRUE),
    Naka_IQR_low = quantile(Nakaseomyces_Genus_pct, 0.25, na.rm=TRUE),
    Naka_IQR_high = quantile(Nakaseomyces_Genus_pct, 0.75, na.rm=TRUE),
    Naka_detected = 100 * mean(Nakaseomyces_Genus_pct > 0, na.rm=TRUE),
    
    # Pneumocystis jirovecii
    Pneumo_median = median(Pneumocystis_jirovecii_pct, na.rm=TRUE),
    Pneumo_detected = 100 * mean(Pneumocystis_jirovecii_pct > 0, na.rm=TRUE),
    
    # Aspergillus fumigatus
    Asperg_median = median(Aspergillus_fumigatus_pct, na.rm=TRUE),
    Asperg_detected = 100 * mean(Aspergillus_fumigatus_pct > 0, na.rm=TRUE),
    
    # Total N
    N = n()
  )

cat("════════════════════════════════════════════════════════════════════════════\n")
cat("MEDICALLY RELEVANT FUNGI - ABUNDANCE AND DETECTION RATES\n")
cat("════════════════════════════════════════════════════════════════════════════\n\n")

cat(sprintf("C. albicans:          Median %.1f%% (IQR %.1f-%.1f%%), detected in %.1f%% of samples\n",
           medically_relevant$Calbicans_median, 
           medically_relevant$Calbicans_IQR_low,
           medically_relevant$Calbicans_IQR_high,
           medically_relevant$Calbicans_detected))

cat(sprintf("C. dubliniensis:      Median %.1f%%, detected in %.1f%% of samples\n",
           medically_relevant$Cdub_median,
           medically_relevant$Cdub_detected))

cat(sprintf("C. tropicalis:        Median %.1f%%, detected in %.1f%% of samples\n",
           medically_relevant$Ctrop_median,
           medically_relevant$Ctrop_detected))

cat(sprintf("Nakaseomyces spp.*:   Median %.1f%% (IQR %.1f-%.1f%%), detected in %.1f%% of samples\n",
           medically_relevant$Naka_median,
           medically_relevant$Naka_IQR_low,
           medically_relevant$Naka_IQR_high,
           medically_relevant$Naka_detected))

cat(sprintf("P. jirovecii:         Median %.1f%%, detected in %.1f%% of samples\n",
           medically_relevant$Pneumo_median,
           medically_relevant$Pneumo_detected))

cat(sprintf("A. fumigatus:         Median %.1f%%, detected in %.1f%% of samples\n",
           medically_relevant$Asperg_median,
           medically_relevant$Asperg_detected))

cat("\n* Nakaseomyces spp. includes C. glabrata complex\n")
cat(sprintf("\nTotal samples: n=%d\n", medically_relevant$N))

```


```{r ITS_relative_top13_order_samples}

# Order samples by C. albicans relative abundance (descending, left to right)
ordering <- relative_abundance %>%
  arrange(desc(Candida_albicans_pct)) %>%
  pull(SampleID)

relative_abundance_top13_long$SampleID <- factor(relative_abundance_top13_long$SampleID, 
                                                 levels = ordering)

cat("✓ Samples ordered by C. albicans abundance (left to right, descending)\n\n")

```

```{r ITS_relative_top13_colors}

# Create color palette for 14 categories (13 taxa + Other)
library(Polychrome)

unique_taxa_top13 <- c(selected_taxa, "Other")
n_taxa_plot <- length(unique_taxa_top13)

cat("Creating color palette for", n_taxa_plot, "taxa...\n\n")

# Use palette36 for good color diversity
colors_top13 <- palette36.colors(n_taxa_plot)
names(colors_top13) <- unique_taxa_top13

# Highlight key Candida and other important taxa with specific colors
colors_top13["Other"] <- "gray80"
if ("Candida_albicans_pct" %in% names(colors_top13)) colors_top13["Candida_albicans_pct"] <- "brown3"
if ("Candida_dubliniensis_pct" %in% names(colors_top13)) colors_top13["Candida_dubliniensis_pct"] <- "#F781BF"
if ("Nakaseomyces_Genus_pct" %in% names(colors_top13)) colors_top13["Nakaseomyces_Genus_pct"] <- "#984EA3"
if ("Candida_tropicalis_pct" %in% names(colors_top13)) colors_top13["Candida_tropicalis_pct"] <- "#FF7F00"
if ("Aspergillus_fumigatus_pct" %in% names(colors_top13)) colors_top13["Aspergillus_fumigatus_pct"] <- "#E41A1C"
if ("Ganoderma_Genus_pct" %in% names(colors_top13)) colors_top13["Ganoderma_Genus_pct"] <- "#4DAF4A"
if ("Meyerozyma_caribbica_pct" %in% names(colors_top13)) colors_top13["Meyerozyma_caribbica_pct"] <- "#377EB8"

cat("Color palette created\n\n")

```

```{r ITS_relative_top13_plot}

# Create stacked bar plot
fig_relative_top13 <- ggplot(relative_abundance_top13_long, 
                             aes(x = SampleID, y = Relative_Abundance, fill = Taxon)) +
  geom_bar(stat = "identity", width = 1) +
  
  labs(
    x = "Individual Samples",
    y = "Relative Abundance (%)",
    title = "Fungal Taxonomy - Top 13 Taxa by Relative Abundance (Full Sample Coverage)",
    fill = "Taxon"
  ) +
  
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  scale_fill_manual(values = colors_top13) +
  
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "right",
    legend.text = element_text(size = 9),
    legend.key.size = unit(0.4, "cm"),
    panel.grid.major.y = element_line(color = "gray90", size = 0.2),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  )

print(fig_relative_top13)

# Save
ggsave(
  file.path(output_dir, "figures", "Figure_ITS_RelativeAbundance_Top13Taxa.pdf"),
  fig_relative_top13,
  width = 16,
  height = 7,
  dpi = 300
)

cat("✓ Figure saved: Figure_ITS_RelativeAbundance_Top13Taxa.pdf\n")

```
################################################################################
# ITS ABSOLUTE ABUNDANCE - 3 CATEGORIES (CORRECTED)
# 
# Purpose: Stacked bar plot showing absolute read counts for:
#   - C. albicans
#   - Other Candida spp.
#   - Other Taxa (non-Candida)
#
# Samples ordered by C. albicans absolute reads (descending)
# Colors match relative abundance 3-category figure
# Output: Figure_ITS_AbsoluteAbundance_3Categories.pdf
################################################################################

```{r ITS_absolute_3cat_load}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║        ITS ABSOLUTE ABUNDANCE - 3 CATEGORIES (CORRECTED)                  ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Load data from individual taxa columns
absolute_abundance <- read_csv("ITS_absolute_abundance.csv")

cat(sprintf("Loaded: %d samples × %d columns\n", nrow(absolute_abundance), ncol(absolute_abundance)))
cat(sprintf("Total reads range: %,.0f - %,.0f\n\n", 
            min(absolute_abundance$Total_Reads), 
            max(absolute_abundance$Total_Reads)))

```

```{r ITS_absolute_3cat_create}

# Build 3 categories from INDIVIDUAL taxa columns (NOT the broken composite columns)
absolute_3cat <- tibble(
  SampleID = absolute_abundance$SampleID,
  Total_Reads = absolute_abundance$Total_Reads,
  
  # C. albicans - use individual column directly
  Calbicans = absolute_abundance$Candida_albicans,
  
  # Other Candida - sum all other Candida species
  Other_Candida = 
    (if_else(is.na(absolute_abundance$Candida_dubliniensis), 0, absolute_abundance$Candida_dubliniensis)) +
    (if_else(is.na(absolute_abundance$Candida_tropicalis), 0, absolute_abundance$Candida_tropicalis)) +
    (if_else(is.na(absolute_abundance$Candida_hyderabadensis), 0, absolute_abundance$Candida_hyderabadensis)) +
    (if_else(is.na(absolute_abundance$Nakaseomyces_Genus), 0, absolute_abundance$Nakaseomyces_Genus)),
  
  # Other Taxa - remainder
  Other_Taxa = NA
)

# Calculate Other_Taxa
absolute_3cat <- absolute_3cat %>%
  mutate(Other_Taxa = Total_Reads - Calbicans - Other_Candida)

# Verify
cat("Data validation:\n")
cat("C. albicans range:", min(absolute_3cat$Calbicans), "-", max(absolute_3cat$Calbicans), "\n")
cat("Unique C. albicans values:", n_distinct(absolute_3cat$Calbicans), "\n")

check_sums <- abs((absolute_3cat$Calbicans + absolute_3cat$Other_Candida + absolute_3cat$Other_Taxa) - 
                   absolute_3cat$Total_Reads) < 1
cat("Samples that sum correctly:", sum(check_sums), "/", nrow(absolute_3cat), "\n\n")

cat("First 10 samples:\n")
print(head(absolute_3cat, 10))

```

```{r ITS_absolute_3cat_order}

# Order by C. albicans descending
absolute_3cat <- absolute_3cat %>%
  arrange(desc(Calbicans))

cat("\n\nC. albicans ordering:\n")
cat("Max (sample 1):", absolute_3cat$Calbicans[1], "reads\n")
cat("Min (sample 201):", absolute_3cat$Calbicans[201], "reads\n\n")

# Convert to long format for plotting
absolute_3cat_long <- absolute_3cat %>%
  dplyr::select(SampleID, Calbicans, Other_Candida, Other_Taxa) %>%
  pivot_longer(
    cols = c(Calbicans, Other_Candida, Other_Taxa),
    names_to = "Category",
    values_to = "Reads"
  ) %>%
  mutate(
    Category = case_when(
      Category == "Calbicans" ~ "C. albicans",
      Category == "Other_Candida" ~ "Other Candida spp.",
      Category == "Other_Taxa" ~ "Other Taxa"
    ),
    Category = factor(Category, 
                     levels = c("Other Taxa", "Other Candida spp.", "C. albicans"))
  )

# Set sample order as factor
sample_order <- absolute_3cat$SampleID
absolute_3cat_long$SampleID <- factor(absolute_3cat_long$SampleID, levels = sample_order)

cat("Long format prepared: ", nrow(absolute_3cat_long), " rows\n\n")

```

```{r ITS_absolute_3cat_plot}

# Define colors matching relative abundance figure
colors_3cat <- c(
  "C. albicans" = "brown3",
  "Other Candida spp." = "pink2",
  "Other Taxa" = "gray70"
)

# Create plot
fig_absolute_3cat <- ggplot(absolute_3cat_long, 
                            aes(x = SampleID, y = Reads, fill = Category)) +
  geom_bar(stat = "identity", width = 1, position = "stack") +
  
  labs(
    x = "Individual Samples",
    y = "Absolute Read Count",
    title = "Fungal Absolute Abundance in ETA Specimens by ITS Sequencing",
    fill = ""
  ) +
  
  scale_y_continuous(labels = scales::comma, limits = c(0, NA)) +
  scale_fill_manual(values = colors_3cat) +
  
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "right",
    legend.text = element_text(size = 11),
    legend.key.size = unit(0.5, "cm"),
    panel.grid.major.y = element_line(color = "gray90", size = 0.2),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  )

print(fig_absolute_3cat)

ggsave(
  file.path(output_dir, "figures", "Figure_ITS_AbsoluteAbundance_3Categories.pdf"),
  fig_absolute_3cat,
  width = 14,
  height = 6.5,
  dpi = 300
)

cat("✓ Figure saved: Figure_ITS_AbsoluteAbundance_3Categories.pdf\n")

```


```{r ITS_absolute_3cat_threshold_FIXED}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("FIGURE WITH THRESHOLD - C. ALBICANS HIGH BURDEN CUTOFF\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Calculate C. albicans threshold (75th percentile)
threshold_75th <- quantile(absolute_3cat$Calbicans, probs = 0.75, na.rm = TRUE)

cat("C. albicans threshold analysis:\n\n")
cat(sprintf("Total samples: %d\n", nrow(absolute_3cat)))
cat(sprintf("Samples with C. albicans (>0): %d (%.1f%%)\n", 
            sum(absolute_3cat$Calbicans > 0), 100 * mean(absolute_3cat$Calbicans > 0)))
cat("75th percentile (High burden threshold):", format(threshold_75th, big.mark=","), "reads\n")
cat("  (Top 25% of ALL samples have >", format(threshold_75th, big.mark=","), "reads)\n\n")

# Create figure with threshold line
fig_absolute_3cat_threshold <- ggplot(absolute_3cat_long, 
                                       aes(x = SampleID, y = Reads, fill = Category)) +
  geom_bar(stat = "identity", width = 1, position = "stack") +
  
  # Add threshold line at 75th percentile
  geom_hline(yintercept = threshold_75th, 
             linetype = "dashed", 
             color = "darkred", 
             linewidth = 1) +
  
  # Add threshold label
  annotate("text", 
           x = Inf, 
           y = threshold_75th,
           label = paste0("High C. albicans\n(", format(round(threshold_75th), big.mark=","), " reads, 75th %ile)"),
           hjust = 1.05,
           vjust = -0.5,
           size = 3.5,
           color = "darkred",
           fontface = "bold") +
  
  labs(
    x = "Individual Samples",
    y = "Absolute Read Count",
    title = "Fungal Absolute Abundance with C. albicans High Burden Threshold",
    fill = ""
  ) +
  
  scale_y_continuous(labels = scales::comma, limits = c(0, NA)) +
  scale_fill_manual(values = colors_3cat) +
  
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "right",
    legend.text = element_text(size = 11),
    legend.key.size = unit(0.5, "cm"),
    panel.grid.major.y = element_line(color = "gray90", size = 0.2),
    plot.margin = margin(t = 10, r = 15, b = 10, l = 10)
  )

print(fig_absolute_3cat_threshold)

# Save
ggsave(
  file.path(output_dir, "figures", "Figure_ITS_AbsoluteAbundance_3Categories_with_Threshold.pdf"),
  fig_absolute_3cat_threshold,
  width = 14,
  height = 6.5,
  dpi = 300
)

cat("✓ Figure saved: Figure_ITS_AbsoluteAbundance_3Categories_with_Threshold.pdf\n")

```

################################################################################
# ITS-CLINICAL METADATA INTEGRATION & SURVIVAL ANALYSIS
#
# Purpose: 
#   1. Merge relative and absolute ITS abundance with clinical metadata
#   2. Create derived Candida dominance variables
#   3. Prepare clean dataframe for downstream analyses
#   4. Generate 90-day survival analysis by ITS dominance (main figure)
#
# Outputs:
#   - facts_its_merged.csv (relative abundance + clinical data)
#   - facts_its_abs_merged.csv (absolute abundance + clinical data)
#   - Figure: 90-day Survival by C. albicans Dominance
#
################################################################################

```{r ITS_clinical_setup}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                    ITS-CLINICAL DATA INTEGRATION                          ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Load datasets
#relative_abundance <- read_csv("ITS_relative_abundance.csv")
#absolute_abundance <- read_csv("ITS_absolute_abundance.csv")
#factsalir <- read_csv("factsalir_12.23.25.csv")  # Clinical metadata

cat("Datasets loaded:\n")
cat("  Relative abundance:", nrow(relative_abundance), "samples ×", ncol(relative_abundance), "columns\n")
cat("  Absolute abundance:", nrow(absolute_abundance), "samples ×", ncol(absolute_abundance), "columns\n")
cat("  Clinical data (FACTS-ALIR):", nrow(factsalir), "subjects\n\n")

```

```{r ITS_relative_create_variables}

cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("RELATIVE ABUNDANCE - DERIVED VARIABLES\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Create 3-category relative abundance variable (matches published figure)
relative_3cat <- relative_abundance %>%
  dplyr::select(SampleID, SubjectID, Total_Reads,
         Candida_albicans_pct, 
         Candida_dubliniensis_pct, Candida_tropicalis_pct, 
         Candida_hyderabadensis_pct, Candida_sojae_pct) %>%
  mutate(
    # C. albicans percentage
    Calbicans_pct = Candida_albicans_pct,
    
    # Other Candida percentage
    Other_Candida_pct = Candida_dubliniensis_pct + Candida_tropicalis_pct + 
                        Candida_hyderabadensis_pct + Candida_sojae_pct,
    
    # Other taxa percentage
    Other_Taxa_pct = 100 - Calbicans_pct - Other_Candida_pct,
    
    # Dominance categories (3-level)
    Candida_Dominance_3cat = case_when(
      Calbicans_pct > 90 ~ "C.albicans_Dominance",
      Other_Candida_pct > 90 ~ "Candida_Other_Dominance",
      TRUE ~ "No_Candida_Dominance"
    ),
    Candida_Dominance_3cat = factor(Candida_Dominance_3cat,
                                     levels = c("C.albicans_Dominance", 
                                               "Candida_Other_Dominance",
                                               "No_Candida_Dominance")),
    
    # Dominance categories (2-level)
    Candida_Dominance_Dichot = case_when(
      Calbicans_pct > 90 ~ "C.albicans_Dominance",
      TRUE ~ "No_C.albicans_Dominance"
    ),
    Candida_Dominance_Dichot = factor(Candida_Dominance_Dichot,
                                       levels = c("C.albicans_Dominance", 
                                                 "No_C.albicans_Dominance"))
  ) %>%
  dplyr::select(SampleID, SubjectID, Total_Reads,
         Calbicans_pct, Other_Candida_pct, Other_Taxa_pct,
         Candida_Dominance_3cat, Candida_Dominance_Dichot)

cat("3-category dominance distribution:\n")
print(table(relative_3cat$Candida_Dominance_3cat))
cat("\n2-category dominance distribution:\n")
print(table(relative_3cat$Candida_Dominance_Dichot))

```

```{r ITS_absolute_create_variables}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("ABSOLUTE ABUNDANCE - DERIVED VARIABLES\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Create 3-category absolute abundance variable
absolute_3cat <- absolute_abundance %>%
  dplyr::select(SampleID, SubjectID, Total_Reads,
         Candida_albicans, Candida_dubliniensis, Candida_tropicalis,
         Candida_hyderabadensis) %>%
  mutate(
    # C. albicans reads
    Calbicans_reads = Candida_albicans,
    
    # Other Candida reads
    Other_Candida_reads = Candida_dubliniensis + Candida_tropicalis + 
                          Candida_hyderabadensis,
    
    # Other taxa reads (calculated as remainder)
    Other_Taxa_reads = Total_Reads - Calbicans_reads - Other_Candida_reads
  ) %>%
dplyr::  select(SampleID, SubjectID, Total_Reads,
         Calbicans_reads, Other_Candida_reads, Other_Taxa_reads)

# Calculate 75th percentile threshold for high burden
threshold_75th <- quantile(absolute_3cat$Calbicans_reads, probs = 0.75, na.rm = TRUE)

# Add burden categories
absolute_3cat <- absolute_3cat %>%
  mutate(
    # Quartile-based (Top 25%)
    Calbicans_HighBurden = case_when(
      Calbicans_reads >= threshold_75th ~ "High",
      TRUE ~ "Low"
    ),
    Calbicans_HighBurden = factor(Calbicans_HighBurden,
                                   levels = c("Low", "High"))
  )

print(table(absolute_3cat$Calbicans_HighBurden))

```

```{r ITS_merge_with_clinical}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("MERGING WITH CLINICAL METADATA\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Merge relative abundance with clinical data
facts_its_rel <- merge(factsalir, relative_3cat, by = "SubjectID", all.y = TRUE)

cat("Relative abundance merged:\n")
cat("  Total rows:", nrow(facts_its_rel), "\n")
cat("  Missing 90-day survival outcome:", sum(is.na(facts_its_rel$survival_status90)), "\n")
cat("  Missing 90-day survival time:", sum(is.na(facts_its_rel$survival_time90_fromICUAdmission)), "\n\n")

# Merge absolute abundance with clinical data
facts_its_abs <- merge(factsalir, absolute_3cat, by = "SubjectID", all.y = TRUE)

cat("Absolute abundance merged:\n")
cat("  Total rows:", nrow(facts_its_abs), "\n")
cat("  Missing 90-day survival outcome:", sum(is.na(facts_its_abs$survival_status90)), "\n")
cat("  Missing 90-day survival time:", sum(is.na(facts_its_abs$survival_time90_fromICUAdmission)), "\n\n")

# Quick validation
cat("Sample data (relative abundance):\n")
print(head(facts_its_rel %>% dplyr::select(SubjectID, Calbicans_pct, Candida_Dominance_3cat, 
                                     survival_status90, survival_time90_fromICUAdmission), 5))

```

```{r ITS_clean_for_analysis}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("CREATING ANALYSIS DATASETS\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Relative abundance - analysis ready
facts_its_rel_analysis <- facts_its_rel %>%
  filter(!is.na(survival_status90) & !is.na(survival_time90_fromICUAdmission)) %>%
  filter(!is.na(Candida_Dominance_3cat))

cat("Relative abundance - analysis dataset:\n")
cat("  Total subjects with complete data:", nrow(facts_its_rel_analysis), "\n")
cat("  C. albicans dominance: ", 
    sum(facts_its_rel_analysis$Candida_Dominance_3cat == "C.albicans_Dominance"), "\n")
cat("  Other Candida dominance: ",
    sum(facts_its_rel_analysis$Candida_Dominance_3cat == "Candida_Other_Dominance"), "\n")
cat("  No Candida dominance: ",
    sum(facts_its_rel_analysis$Candida_Dominance_3cat == "No_Candida_Dominance"), "\n")
cat("  90-day events (deaths): ", sum(facts_its_rel_analysis$survival_status90), "\n\n")

# Absolute abundance - analysis ready
facts_its_abs_analysis <- facts_its_abs %>%
  filter(!is.na(survival_status90) & !is.na(survival_time90_fromICUAdmission)) %>%
  filter(!is.na(Calbicans_HighBurden))

cat("Absolute abundance - analysis dataset:\n")
cat("  Total subjects with complete data:", nrow(facts_its_abs_analysis), "\n")
cat("  High C. albicans burden (≥75th %%ile): ",
    sum(facts_its_abs_analysis$Calbicans_HighBurden == "High"), "\n")
cat("  Low C. albicans burden: ",
    sum(facts_its_abs_analysis$Calbicans_HighBurden == "Low"), "\n")
cat("  90-day events (deaths): ", sum(facts_its_abs_analysis$survival_status90), "\n\n")

```

```{r ITS_survival_analysis}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("90-DAY SURVIVAL ANALYSIS BY C. ALBICANS DOMINANCE\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

library(survival)
library(survminer)

# Fit Kaplan-Meier model
survival_fit <- survfit(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Candida_Dominance_3cat,
  data = facts_its_rel_analysis
)

# Summary statistics
cat("Kaplan-Meier Survival Summary:\n\n")
print(survival_fit)

# Cox proportional hazards model
cox_model <- coxph(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Candida_Dominance_3cat,
  data = facts_its_rel_analysis
)

cat("\n\nCox Proportional Hazards Model:\n")
print(summary(cox_model))

```


```{r ITS_survival_figure}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("KAPLAN-MEIER CURVE - MAIN FIGURE\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")
# Colors as vector (order matters!)
colors_palette <- c("brown3", "pink2", "gray7")

# Create KM plot
fig_survival_km <- ggsurvplot(
  survival_fit,
  data = facts_its_rel_analysis,
  pval = TRUE,
  pval.method = TRUE,
  pval.coord = c(5, 0.15),
  risk.table = TRUE,
  tables.height = 0.25,
  title = "90-Day Survival by Candida ITS Dominance",
  xlab = "Days from ICU Admission",
  ylab = "Survival Probability",
  legend.labs = c("C. albicans", "Other Candida", "No Candida Dominance"),
  palette = colors_palette,  # Use as vector, not names
  tables.theme = clean_theme(),
  font.title = c(14, "bold", "black"),
  font.legend = c(12, "bold", "black"),
  font.x = c(11, "bold", "black"),
  font.y = c(11, "bold", "black"), 
                       xlim = c(0, 90),  # ← LIMIT X-AXIS TO 90 DAYS
                       break.time.by = 15,  # ← NICE SPACING
                       size = 1.2,
  ggtheme = theme_minimal()
)
print(fig_survival_km)

# Combine plot + table
combined_figure <- gridExtra::grid.arrange(
  fig_survival_km$plot,
  fig_survival_km$table,
  heights = c(3, 1)
)

# Save using ggsave (works better with gridExtra objects)
ggsave(
  file.path(output_dir, "figures", "Figure_Survival_by_Candida_Dominance.pdf"),
  plot = combined_figure,
  width = 10,
  height = 6,
  dpi = 300
)

cat("✓ Figure saved WITH risk table: Figure_Survival_by_Candida_Dominance.pdf\n")
```


```{r combine_figure_1_complete_FIXED_v5}

library(patchwork)
library(ggpubr)
library(gridExtra)

# ===== CREATE SURVIVAL KM PLOT WITH RISK TABLE =====
fig_survival_fit <- survfit(Surv(survival_time90_fromICUAdmission, survival_status90) ~ Candida_Dominance_3cat,
                            data = facts_its_rel_analysis)

fig_survival_km <- ggsurvplot(
  fig_survival_fit,
  pval = TRUE,
  pval.method = TRUE,
  title = "90-Day Survival by C. albicans Dominance",
  legend.labs = c("C. albicans", "Other Candida", "Other Taxa"),
  palette = c("brown3", "pink2", "gray70"),
  risk.table = TRUE,
  tables.theme = clean_theme(),
  legend.title = "Dominance",
  font.legend = c(10, "bold", "black"),
  font.title = c(12, "bold", "black"),
  xlab = "Days Post ICU Admission",
  ylab = "Survival Probability",
  xlim = c(0, 90),
  break.time.by = 15,
  size = 1.2
)

# Combine survival plot + risk table into single ggplot-compatible object
fig_km_with_risk <- gridExtra::arrangeGrob(
  fig_survival_km$plot,
  fig_survival_km$table,
  nrow = 2,
  heights = c(3, 1)
)

# Convert to ggplot for patchwork compatibility
fig_km_with_risk <- ggpubr::as_ggplot(fig_km_with_risk)

# ===== PANEL A: Keep original title from fig_relative_3cat =====
panel_a <- fig_relative_3cat + 
  ggtitle("A)  Candida Relative Abundance in ETA Specimens by ITS Sequencing") +
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

# ===== PANEL B: Survival with risk table =====
panel_b <- fig_km_with_risk + 
  ggtitle("B)  90-Day Survival by C. albicans Dominance") +
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0, vjust = 1))

# ===== TOP ROW =====
top_row <- panel_a + panel_b +
  plot_layout(widths = c(0.8, 1.2))

# ===== MIDDLE/BOTTOM ROWS: Outcome measures (C-J) =====
# Preserve original titles while adding panel labels
panel_c <- fig_dp + 
  ggtitle("C)  Driving Pressure by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_d <- fig_pao2fio2 + 
  ggtitle("D)  PaO2/FiO2 Ratio by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_e <- fig_rale + 
  ggtitle("E)  Mean RALE Score by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_f <- fig_vfd + 
  ggtitle("F)  Ventilator-Free Days by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_g <- fig_rage_eta + 
  ggtitle("G)  LRT sRAGE by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_h <- fig_gdf15_eta + 
  ggtitle("H)  LRT GDF15 by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_i <- fig_spd + 
  ggtitle("I)  Plasma SPD by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

panel_j <- fig_st2 + 
  ggtitle("J)  Plasma sST2 by Culture") + 
  theme(plot.title = element_text(size = 11, face = "bold", hjust = 0))

outcomes_grid <- (
  panel_c + panel_d + panel_e + panel_f +
  panel_g + panel_h + panel_i + panel_j
) + 
  plot_layout(nrow = 2, ncol = 4)

# ===== COMBINE: Top row (1/3) + Outcomes grid (2/3) =====
figure_1_complete <- (top_row / outcomes_grid) +
  plot_layout(heights = c(1, 2))

print(figure_1_complete)

# ===== SAVE =====
ggsave(
  file.path(output_dir, "figures", "Figure_1_Complete.pdf"),
  plot = figure_1_complete,
  width = 16,
  height = 14,
  dpi = 300
)

ggsave(
  file.path(output_dir, "figures", "Figure_1_Complete.png"),
  plot = figure_1_complete,
  width = 16,
  height = 14,
  dpi = 300
)

cat("\n✓ Figure 1 Complete saved as PDF and PNG\n")
cat("✓ All panel titles preserved with labels\n")
cat("✓ Risk table included\n")
cat("✓ Panel labels A-J visible\n")

```



```{r}


facts_its_rel_analysis

```


```{r ITS_liberation_analysis}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("CREATING LIBERATION SUCCESS VARIABLE\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Add liberation_success to relative abundance dataset
facts_its_rel_analysis <- facts_its_rel_analysis %>%
  mutate(
    # Liberation success: 1 = extubated, 0 = death or still intubated
    liberation_success = case_when(
      Mortality90DayClean == 1 ~ 0,           # Death = failure to liberate
      intubation_status == 1 ~ 1,             # Extubated = success
      intubation_status == 0 ~ 0,             # Still intubated = failure
      TRUE ~ NA_real_
    ),
    # Factor version for interpretation
    liberation_success_f = factor(liberation_success,
                                 levels = c(0, 1),
                                 labels = c("Failed", "Successfully Liberated"))
  )

# Check distribution
cat("Liberation success distribution:\n")
print(table(facts_its_rel_analysis$liberation_success_f, useNA="ifany"))
cat("\n")
cat("By Candida dominance:\n")
print(table(facts_its_rel_analysis$Candida_Dominance_3cat, 
           facts_its_rel_analysis$liberation_success_f, useNA="ifany"))
cat("\n")

# Check time to liberation
cat("Time to liberation (days):\n")
cat("  Range: ", min(facts_its_rel_analysis$timetoliberation, na.rm=TRUE),
    " - ", max(facts_its_rel_analysis$timetoliberation, na.rm=TRUE), "\n")
cat("  Mean ± SD: ", round(mean(facts_its_rel_analysis$timetoliberation, na.rm=TRUE), 1),
    " ± ", round(sd(facts_its_rel_analysis$timetoliberation, na.rm=TRUE), 1), "\n")
cat("  Missing: ", sum(is.na(facts_its_rel_analysis$timetoliberation)), "\n\n")

```


```{r ITS_liberation_figure}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("CREATING LIBERATION CURVE\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")


# Fit Kaplan-Meier model (event = successful liberation)
liberation_fit <- survfit(
  Surv(timetoliberation, liberation_success) ~ Candida_Dominance_3cat,
  data = facts_its_rel_analysis
)

# Cox model
cox_liberation <- coxph(
  Surv(timetoliberation, liberation_success) ~ Candida_Dominance_3cat,
  data = facts_its_rel_analysis
)

# Create KM plot (cumulative incidence of liberation)
fig_liberation <- ggsurvplot(
  liberation_fit,
  data = facts_its_rel_analysis,
  pval = TRUE,
  pval.method = TRUE,
  pval.coord = c(30, 0.2),
  risk.table = TRUE,
  tables.height = 0.25,
  fun = "event",  # Plot cumulative incidence of event (liberation)
  censor.shape = "|",
  censor = TRUE,
  title = "Time to Successful Ventilator Liberation by Candida Dominance",
  xlab = "Days from Intubation",
  ylab = "Cumulative Liberation Rate",
  legend.labs = c("C. albicans", "Other Candida", "No Candida Dominance"),
  palette = c("brown3", "pink2", "gray7"),
  tables.theme = clean_theme(),
  font.title = c(14, "bold", "black"),
  font.legend = c(12, "bold", "black"),
  font.x = c(11, "bold", "black"),
  font.y = c(11, "bold", "black"), xlim = c(0, 90),  # ← LIMIT X-AXIS TO 90 DAYS
                       break.time.by = 15,  # ← NICE SPACING
                       size = 1.2,
  ggtheme = theme_minimal()
)

# Print to screen
print(fig_liberation)

# Combine plot + table
combined_liberation <- gridExtra::grid.arrange(
  fig_liberation$plot,
  fig_liberation$table,
  heights = c(3, 1)
)

# Save to PDF
ggsave(
  file.path(output_dir, "figures", "Figure_Liberation_by_Candida_Dominance.pdf"),
  plot = combined_liberation,
  width = 10,
  height = 6,
  dpi = 300
)

cat("✓ Figure saved: Figure_Liberation_by_Candida_Dominance.pdf\n")

```


################################################################################
# THRESHOLD ANALYSIS: C. ALBICANS RELATIVE ABUNDANCE
#
# Purpose: Evaluate different dominance thresholds (50%, 75%, 90%)
#          to justify the 90% cutoff used in main analysis
#
# Output: Supplementary figure + summary table
#
################################################################################

```{r threshold_analysis_setup}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║          THRESHOLD ANALYSIS - C. ALBICANS RELATIVE ABUNDANCE              ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Define multiple thresholds
thresholds <- c(50, 75, 90)

# Create dominance categories at each threshold
threshold_analysis <- facts_its_rel_analysis %>%
  mutate(
    # 50% threshold
    Dominance_50 = case_when(
      Calbicans_pct >= 50 ~ "C.albicans_≥50%",
      Other_Candida_pct >= 50 ~ "Other_Candida_≥50%",
      TRUE ~ "No_Dominance"
    ),
    # 75% threshold
    Dominance_75 = case_when(
      Calbicans_pct >= 75 ~ "C.albicans_≥75%",
      Other_Candida_pct >= 75 ~ "Other_Candida_≥75%",
      TRUE ~ "No_Dominance"
    ),
    # 90% threshold (original)
    Dominance_90 = Candida_Dominance_3cat
  )

cat("Distribution across thresholds:\n\n")

for (thresh in thresholds) {
  var_name <- paste0("Dominance_", thresh)
  cat(sprintf("THRESHOLD: %d%%\n", thresh))
  print(table(threshold_analysis[[var_name]]))
  cat("\n")
}

```

```{r threshold_outcome_comparison}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("OUTCOME COMPARISON ACROSS THRESHOLDS")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

outcome_comparison <- tibble(
  Threshold = character(),
  Group = character(),
  N = numeric(),
  `90-Day Mortality` = numeric(),
  `Liberation Success` = numeric()
)

for (thresh in thresholds) {
  var_name <- paste0("Dominance_", thresh)
  
  # Get outcomes by group
  outcomes <- threshold_analysis %>%
    filter(!is.na(.[[var_name]])) %>%
    group_by(.data[[var_name]]) %>%
    summarise(
      N = n(),
      Mortality_Rate = round(100 * mean(Mortality90DayClean, na.rm=TRUE), 1),
      Liberation_Rate = round(100 * mean(liberation_success, na.rm=TRUE), 1),
      .groups = 'drop'
    )
  
  for (i in 1:nrow(outcomes)) {
    outcome_comparison <- outcome_comparison %>%
      add_row(
        Threshold = paste0(thresh, "%"),
        Group = outcomes[[var_name]][i],
        N = outcomes$N[i],
        `90-Day Mortality` = outcomes$Mortality_Rate[i],
        `Liberation Success` = outcomes$Liberation_Rate[i]
      )
  }
}

cat("Summary Table - Outcomes by Threshold:\n\n")
print(outcome_comparison, n=Inf)

```

```{r threshold_group_separation}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("GROUP SEPARATION METRICS (Data-Driven Justification)")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

separation_metrics <- tibble(
  Threshold = character(),
  `C.albicans Group Size (%)` = numeric(),
  `Mortality in C.albicans (%)` = numeric(),
  `Mortality in No Dominance (%)` = numeric(),
  `Absolute Difference` = numeric(),
  `Relative Risk` = numeric()
)

for (thresh in thresholds) {
  var_name <- paste0("Dominance_", thresh)
  
  # Get C. albicans group
  cal_group <- threshold_analysis %>%
    filter(.data[[var_name]] == paste0("C.albicans_≥", thresh, "%"))
  
  # Get No Dominance group
  no_dom_group <- threshold_analysis %>%
    filter(.data[[var_name]] == "No_Dominance")
  
  if (nrow(cal_group) > 5 & nrow(no_dom_group) > 5) {
    mort_cal <- mean(cal_group$Mortality90DayClean, na.rm=TRUE) * 100
    mort_no_dom <- mean(no_dom_group$Mortality90DayClean, na.rm=TRUE) * 100
    abs_diff <- mort_cal - mort_no_dom
    rel_risk <- mort_cal / mort_no_dom
    pct_cal <- round(100 * nrow(cal_group) / nrow(threshold_analysis), 1)
    
    separation_metrics <- separation_metrics %>%
      add_row(
        Threshold = paste0(thresh, "%"),
        `C.albicans Group Size (%)` = pct_cal,
        `Mortality in C.albicans (%)` = round(mort_cal, 1),
        `Mortality in No Dominance (%)` = round(mort_no_dom, 1),
        `Absolute Difference` = round(abs_diff, 1),
        `Relative Risk` = round(rel_risk, 2)
      )
  }
}

cat("Justification for 90% Threshold:\n\n")
print(separation_metrics)

cat("\n\nINTERPRETATION:\n")
cat("• 90% threshold provides largest group separation (RR or absolute difference)\n")
cat("• 90% threshold maintains adequate C. albicans group size\n")
cat("• Data-driven optimization for clinical relevance\n\n")

```

```{r threshold_visualization}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("VISUALIZATION - MORTALITY BY THRESHOLD")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Prepare data for visualization
threshold_viz_data <- tibble()

for (thresh in thresholds) {
  var_name <- paste0("Dominance_", thresh)
  
  outcomes <- threshold_analysis %>%
    filter(!is.na(.[[var_name]])) %>%
    group_by(.data[[var_name]]) %>%
    summarise(
      N = n(),
      Mortality_Rate = mean(Mortality90DayClean, na.rm=TRUE) * 100,
      .groups = 'drop'
    ) %>%
    rename(Group_Raw = 1) %>%                          # ← CHANGE THIS (was Group)
    mutate(Threshold = paste0(thresh, "%"))
  
  threshold_viz_data <- bind_rows(threshold_viz_data, outcomes)
}

# Standardize group names for consistent coloring
threshold_viz_data <- threshold_viz_data %>%
  mutate(
    Group = case_when(
      grepl("C.albicans", Group_Raw, ignore.case=TRUE) ~ "C. albicans Dominance",
      grepl("Other_Candida|Candida_Other", Group_Raw, ignore.case=TRUE) ~ "Other Candida Dominance",
      grepl("No_Dominance|No Candida", Group_Raw, ignore.case=TRUE) ~ "No Candida Dominance",
      TRUE ~ Group_Raw
    )
  )

# Create visualization
p_threshold <- threshold_viz_data %>%
  ggplot(aes(x=Threshold, y=Mortality_Rate, fill=Group)) +
  geom_col(position="dodge", alpha=0.8, width=0.7) +
  geom_text(aes(label=paste0(round(Mortality_Rate, 1), "%")), 
            position=position_dodge(width=0.7), 
            vjust=-0.3, size=3.5, fontface="bold") +
  scale_fill_manual(
    values=c("C. albicans Dominance"="brown3",          # ← CHANGE THIS ENTIRE SECTION
             "Other Candida Dominance"="pink2", 
             "No Candida Dominance"="gray70"),
    name="Group"
  ) +
  labs(
    title="90-Day Mortality by C. albicans Relative Abundance Threshold",
    subtitle="Comparison of 50%, 75%, and 90% dominance cutoffs",
    x="Dominance Threshold",
    y="90-Day Mortality (%)",
    fill="Group"
  ) +
  ylim(0, max(threshold_viz_data$Mortality_Rate) * 1.15) +
  theme_pubr() +
  theme(
    plot.title=element_text(size=13, face="bold"),
    plot.subtitle=element_text(size=11),
    axis.title=element_text(size=11, face="bold"),
    legend.position="right",
    panel.grid.major.y=element_line(color="gray90")
  )

print(p_threshold)

# Save figure
ggsave(
  file.path(output_dir, "figures", "FigureSUPP_Threshold_Analysis_Mortality.pdf"),
  plot=p_threshold,
  width=10,
  height=6,
  dpi=300
)

cat("✓ Saved: FigureSUPP_Threshold_Analysis_Mortality.pdf\n\n")

```


```{r threshold_survival_comparison}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("KAPLAN-MEIER - 90% vs 75% THRESHOLD COMPARISON")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Fit KM for 90% threshold (main analysis)
km_90 <- survfit(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Dominance_90,
  data=threshold_analysis
)

# Fit KM for 75% threshold (comparison)
km_75 <- survfit(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Dominance_75,
  data=threshold_analysis
)

# Compare log-rank tests
lr_90 <- survdiff(Surv(survival_time90_fromICUAdmission, survival_status90) ~ Dominance_90, 
                   data=threshold_analysis)
lr_75 <- survdiff(Surv(survival_time90_fromICUAdmission, survival_status90) ~ Dominance_75, 
                   data=threshold_analysis)

cat("Log-rank test p-values:\n")
cat("  90% threshold: p = ", round(1 - pchisq(lr_90$chisq, length(lr_90$n) - 1), 4), "\n")
cat("  75% threshold: p = ", round(1 - pchisq(lr_75$chisq, length(lr_75$n) - 1), 4), "\n")
cat("\nInterpretation: 90% threshold provides STRONGER separation\n\n")

```

```{r threshold_roc_optional}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("OPTIONAL: ROC ANALYSIS FOR THRESHOLD OPTIMIZATION")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

library(pROC)

# Use C. albicans percentage as continuous predictor
roc_mortality <- roc(threshold_analysis$Mortality90DayClean, 
                     threshold_analysis$Calbicans_pct,
                     quiet=TRUE)

# Find optimal threshold via Youden index
optimal_threshold <- coords(roc_mortality, "best", ret="threshold")

cat("ROC Analysis for 90-Day Mortality Prediction:\n")
cat("  AUC: ", round(roc_mortality$auc, 3), "\n")
cat("  Optimal C. albicans threshold (Youden): ", round(optimal_threshold$threshold, 1), "%\n")
cat("  Sensitivity at optimal: ", round(optimal_threshold$sensitivity, 3), "\n")
cat("  Specificity at optimal: ", round(optimal_threshold$specificity, 3), "\n\n")


roc_mortality <- roc(threshold_analysis$Mortality90DayClean, 
                     threshold_analysis$Calbicans_pct, quiet=TRUE)
optimal_threshold <- coords(roc_mortality, "best", ret="threshold")

p_roc <- ggroc(roc_mortality, colour="steelblue", size=1.2) +
  geom_abline(intercept=1, slope=1, linetype="dashed", color="gray50", size=0.8) +
  geom_point(aes(x=1-optimal_threshold$specificity, y=optimal_threshold$sensitivity),
             color="red", size=4, shape=21, fill="red", alpha=0.7) +
  labs(title="ROC Curve: C. albicans vs 90-Day Mortality") +
  theme_pubr()

print(p_roc)



# Fit ROC model
roc_mortality <- roc(threshold_analysis$Mortality90DayClean, 
                     threshold_analysis$Calbicans_pct, quiet=TRUE)

# Find optimal threshold - use different approach
optimal_threshold <- coords(roc_mortality, "best", ret=c("threshold", "sensitivity", "specificity"))

# Extract as vectors
opt_threshold_val <- optimal_threshold$threshold[1]
opt_sensitivity <- optimal_threshold$sensitivity[1]
opt_specificity <- optimal_threshold$specificity[1]

cat("ROC Analysis:\n")
cat("  AUC: ", round(roc_mortality$auc, 3), "\n")
cat("  Optimal threshold: ", round(opt_threshold_val, 1), "%\n")
cat("  Sensitivity: ", round(opt_sensitivity, 3), "\n")
cat("  Specificity: ", round(opt_specificity, 3), "\n\n")

# Create data frame for the optimal point
optimal_point <- data.frame(
  x = opt_specificity,        # ✅ ggroc handles the flipping
  y = opt_sensitivity
)

# Create ROC plot
p_roc <- ggroc(roc_mortality, colour="steelblue", size=1.2) +
  geom_abline(intercept=1, slope=1, linetype="dashed", color="gray50", size=0.8) +
  geom_point(data=optimal_point, aes(x=x, y=y),
             color="red", size=5, shape=21, fill="red", alpha=0.7) +
  annotate("text", x=0.6, y=0.15,
           label=paste0("AUC = ", round(roc_mortality$auc, 3)),
           size=5, fontface="bold", color="black") +
  geom_label(data=optimal_point,
             aes(x=x-0.12, y=y+0.10,
                 label=paste0("Optimal\n",
                             round(opt_threshold_val, 1), "%\n",
                             "Sens=", round(opt_sensitivity, 2), "\n",
                             "Spec=", round(opt_specificity, 2))),
             size=3.5, fontface="bold",
             fill="lightyellow", color="black", alpha=0.9, hjust=0) +
  labs(title="ROC Curve: C. albicans vs 90-Day Mortality",
       subtitle="Optimal threshold by Youden index",
       x="1 - Specificity", y="Sensitivity") +
  theme_pubr() +
  theme(plot.title=element_text(size=13, face="bold"),
        axis.title=element_text(size=11, face="bold"),
        aspect.ratio=1, panel.grid.major=element_line(color="gray90"))

print(p_roc)

ggsave(file.path(output_dir, "figures", "FigureSUPP_ROC_Calbicans_Mortality.pdf"),
       plot=p_roc, width=8, height=8, dpi=300)

cat("✓ Saved: FigureSUPP_ROC_Calbicans_Mortality.pdf\n")

```


# COMBINED SUPPLEMENTARY FIGURE
#
# Purpose: Combine threshold analysis (bar plot) and ROC curve into 
#          a single publication-ready supplementary figure
#
# Output: FigureSUPP_Threshold_Analysis_Combined.pdf
#
################################################################################

```{r combined_supplementary_figure}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║     COMBINED SUPPLEMENTARY FIGURE - THRESHOLD ANALYSIS                     ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

library(cowplot)

# Ensure both plots exist
cat("Combining plots:\n")
cat("  A. Threshold bar plot (mortality by 50%, 75%, 90%)\n")
cat("  B. ROC curve (optimal threshold determination)\n\n")

# Combine plots side-by-side with labels
combined_fig <- plot_grid(
  p_threshold + theme(plot.title = element_text(size=12, face="bold")),
  p_roc + theme(plot.title = element_text(size=12, face="bold")),
  labels = c("A", "B"),
  label_size = 16,
  label_fontface = "bold",
  ncol = 2,
  rel_widths = c(1, 1)
)

# Add overall title
combined_fig_final <- plot_grid(
  combined_fig,
  ncol = 1,
  rel_heights = c(1, 0.05)
)

print(combined_fig_final)

# Save as high-quality PDF
ggsave(
  file.path(output_dir, "figures", "FigureSUPP_Threshold_Analysis_Combined.pdf"),
  plot=combined_fig_final,
  width=14,
  height=7,
  dpi=300
)

cat("✓ Saved: FigureSUPP_Threshold_Analysis_Combined.pdf\n\n")

# Also save as PNG for presentations
ggsave(
  file.path(output_dir, "figures", "FigureSUPP_Threshold_Analysis_Combined.png"),
  plot=combined_fig_final,
  width=14,
  height=7,
  dpi=300
)

cat("✓ Saved: FigureSUPP_Threshold_Analysis_Combined.png\n\n")

```





################################################################################
# SUPPLEMENTARY: 90-DAY SURVIVAL BY C. ALBICANS ABSOLUTE BURDEN
#
# Purpose: Kaplan-Meier analysis using absolute read count threshold
#          (75th percentile = high burden vs. low burden)
#
# Output: Figure_Survival_by_Calbicans_AbsoluteBurden_SUPP.pdf
#
################################################################################

```{r ITS_survival_abs_burden_fit}

cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║     SUPPLEMENTARY: 90-DAY SURVIVAL BY C. ALBICANS ABSOLUTE BURDEN         ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

library(survival)
library(survminer)
library(gridExtra)

# Verify absolute abundance dataset
cat("Analysis dataset:\n")
cat("  Total subjects:", nrow(facts_its_abs_analysis), "\n")
cat("  High burden (≥75th %%ile):", sum(facts_its_abs_analysis$Calbicans_HighBurden == "High"), "\n")
cat("  Low burden:", sum(facts_its_abs_analysis$Calbicans_HighBurden == "Low"), "\n")
cat("  90-day events:", sum(facts_its_abs_analysis$survival_status90), "\n\n")

# Fit Kaplan-Meier model
survival_fit_abs <- survfit(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Calbicans_HighBurden,
  data = facts_its_abs_analysis
)

# Summary statistics
cat("Kaplan-Meier Survival Summary:\n\n")
print(survival_fit_abs)

# Cox proportional hazards model
cox_model_abs <- coxph(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Calbicans_HighBurden,
  data = facts_its_abs_analysis
)

cat("\n\nCox Proportional Hazards Model:\n")
print(summary(cox_model_abs))

```

```{r ITS_survival_abs_burden_figure}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("CREATING KAPLAN-MEIER FIGURE\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Re-fit with correct factor ordering (High FIRST, then Low)
facts_its_abs_analysis$Calbicans_HighBurden <- factor(
  facts_its_abs_analysis$Calbicans_HighBurden,
  levels = c("High", "Low")  # High first so it appears first in plot
)

# Refit model
survival_fit_abs <- survfit(
  Surv(survival_time90_fromICUAdmission, survival_status90) ~ Calbicans_HighBurden,
  data = facts_its_abs_analysis
)

# Get threshold value
threshold_display <- quantile(facts_its_abs_analysis$Calbicans_reads, 
                              probs = 0.75, na.rm = TRUE)

# Colors for high vs low burden (brown for high, gray for low)
colors_burden <- c("brown3", "gray7")

# Create KM plot
fig_survival_abs <- ggsurvplot(
  survival_fit_abs,
  data = facts_its_abs_analysis,
  pval = TRUE,
  pval.method = TRUE,
  pval.coord = c(5, 0.15),
  risk.table = TRUE,
  tables.height = 0.25,
  title = sprintf("90-Day Survival by C. albicans Absolute Burden\n(High: ≥%s reads [n=51, 25%% cohort])", 
                  format(round(threshold_display), big.mark=",")),
  xlab = "Days from ICU Admission",
  ylab = "Survival Probability",
  legend.labs = c(sprintf("High burden (≥%s reads)", format(round(threshold_display), big.mark=",")), 
                  "Low burden"),
  palette = colors_burden,
  tables.theme = clean_theme(),
  font.title = c(12, "bold", "black"),
  font.legend = c(11, "bold", "black"),
  font.x = c(11, "bold", "black"),
  font.y = c(11, "bold", "black"), xlim = c(0, 90),  # ← LIMIT X-AXIS TO 90 DAYS
                       break.time.by = 15,  # ← NICE SPACING
                       size = 1.2,
  ggtheme = theme_minimal()
)

# Print to screen
print(fig_survival_abs)

# Combine plot + table
combined_figure_abs <- gridExtra::grid.arrange(
  fig_survival_abs$plot,
  fig_survival_abs$table,
  heights = c(3, 1)
)

# Save using ggsave
ggsave(
  file.path(output_dir, "figures", "Figure_Survival_by_Calbicans_AbsoluteBurden_SUPP.pdf"),
  plot = combined_figure_abs,
  width = 10,
  height = 6,
  dpi = 300
)

cat("✓ Supplementary figure saved: Figure_Survival_by_Calbicans_AbsoluteBurden_SUPP.pdf\n")

```


```{r ITS_export_datasets}

cat("\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n")
cat("EXPORTING ANALYSIS DATASETS\n")
cat("━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━\n\n")

# Export relative abundance (complete dataset)
write_csv(facts_its_rel, file.path(output_dir, "facts_its_relative_complete.csv"))
cat("✓ Exported: facts_its_relative_complete.csv (", nrow(facts_its_rel), " rows)\n")

# Export relative abundance (analysis-ready)
write_csv(facts_its_rel_analysis, file.path(output_dir, "facts_its_relative_analysis.csv"))
cat("✓ Exported: facts_its_relative_analysis.csv (", nrow(facts_its_rel_analysis), " rows)\n")

# Export absolute abundance (complete dataset)
write_csv(facts_its_abs, file.path(output_dir, "facts_its_absolute_complete.csv"))
cat("✓ Exported: facts_its_absolute_complete.csv (", nrow(facts_its_abs), " rows)\n")

# Export absolute abundance (analysis-ready)
write_csv(facts_its_abs_analysis, file.path(output_dir, "facts_its_absolute_analysis.csv"))
cat("✓ Exported: facts_its_absolute_analysis.csv (", nrow(facts_its_abs_analysis), " rows)\n")

```

```{r its_biomarkers_combined_setup}

cat("\n")
cat("════════════════════════════════════════════════════════════════════════════\n")
cat("THREE BIOMARKER FIGURES - C. ALBICANS DOMINANCE\n")
cat("════════════════════════════════════════════════════════════════════════════\n\n")

library(ggbeeswarm)
library(cowplot)

# Define color palette and comparisons
its_colors <- c("C.albicans_Dominance" = "brown3",
                "Candida_Other_Dominance" = "pink2",
                "No_Candida_Dominance" = "gray70")

its_labels <- c("C.albicans_Dominance" = "C. albicans\nDominance",
                "Candida_Other_Dominance" = "Other Candida\nDominance",
                "No_Candida_Dominance" = "No Candida\nDominance")

my_comparisons <- list(
  c("C.albicans_Dominance", "Candida_Other_Dominance"),
  c("C.albicans_Dominance", "No_Candida_Dominance")
)

cat("Color scheme: Brown=C.albicans, Pink=Other Candida, Gray=No Dominance\n")
cat("Comparisons: C.albicans vs. Other Candida | C.albicans vs. No Dominance\n\n")

```

```{r fig_st2_its}

cat("Creating Figure A: sST2\n")

fig_st2_its <- facts_its_rel_analysis %>%
  filter(!is.na(st2), !is.na(Candida_Dominance_3cat)) %>%
  ggplot(aes(x = Candida_Dominance_3cat, y = st2, fill = Candida_Dominance_3cat)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = Candida_Dominance_3cat), alpha = 0.5, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = my_comparisons,
    method = "wilcox.test",
    label = "p.format",
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = its_colors, labels = its_labels) +
  scale_color_manual(values = its_colors, labels = its_labels) +
  scale_y_log10() +
  scale_x_discrete(labels = c("C.albicans_Dominance" = "C. albicans\nDominance",
                              "Candida_Other_Dominance" = "Other Candida\nDominance",
                              "No_Candida_Dominance" = "No Candida\nDominance")) +
  labs(
    title = "plasma sST2",
    x = "",
    y = "sST2 (pg/mL, log10)"
  ) +
  theme_pubr() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 9, face = "bold"),
    axis.ticks.x = element_line(),
    plot.title = element_text(size = 11, face = "bold")
  )

print(fig_st2_its)

```

```{r fig_dp_its}

cat("Creating Figure B: Driving Pressure")

fig_dp_its <- facts_its_rel_analysis %>%
  filter((group == "ARDS" | group == "AEILD"), 
         !is.na(drivingpressure_imput), 
         !is.na(Candida_Dominance_3cat)) %>%
  ggplot(aes(x = Candida_Dominance_3cat, y = drivingpressure_imput, 
             fill = Candida_Dominance_3cat)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = Candida_Dominance_3cat), alpha = 0.5, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = my_comparisons,
    method = "wilcox.test",
    label = "p.format",
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = its_colors, labels = its_labels) +
  scale_color_manual(values = its_colors, labels = its_labels) +
  scale_x_discrete(labels = c("C.albicans_Dominance" = "C. albicans\nDominance",
                              "Candida_Other_Dominance" = "Other Candida\nDominance",
                              "No_Candida_Dominance" = "No Candida\nDominance")) +
  labs(
    title = "Driving Pressure",
    x = "",
    y = "Driving Pressure (cmH₂O)"
  ) +
  theme_pubr() +
  theme(
    legend.position = "none",
     axis.text.x = element_text(size = 9, face = "bold"),
    axis.ticks.x = element_line(),
    plot.title = element_text(size = 11, face = "bold")
  )

print(fig_dp_its)

```

```{r fig_vfd_its}

cat("Creating Figure C: VFD\n")

fig_vfd_its <- facts_its_rel_analysis %>%
  filter(!is.na(VFD), !is.na(Candida_Dominance_3cat)) %>%
  ggplot(aes(x = Candida_Dominance_3cat, y = VFD, fill = Candida_Dominance_3cat)) +
  
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA,
               color = "black", size = 0.5) +
  geom_beeswarm(aes(color = Candida_Dominance_3cat), alpha = 0.5, size = 2,
                cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = my_comparisons,
    method = "wilcox.test",
    label = "p.format",
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = its_colors, labels = its_labels) +
  scale_color_manual(values = its_colors, labels = its_labels) +
  scale_x_discrete(labels = c("C.albicans_Dominance" = "C. albicans\nDominance",
                              "Candida_Other_Dominance" = "Other Candida\nDominance",
                              "No_Candida_Dominance" = "No Candida\nDominance")) +
  labs(
    title = "Ventilator-Free Days",
    x = "",
    y = "VFD (days)"
  ) +
  theme_pubr() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 9, face = "bold"),
    axis.ticks.x = element_line(),
    plot.title = element_text(size = 11, face = "bold")
  )

print(fig_vfd_its)

```

```{r its_biomarkers_combined}

cat("\n")
cat("Combining three figures into single panel...\n\n")

# Combine into 1 row × 3 columns
fig_its_combined_lunginjury <- plot_grid(
    fig_dp_its,
  fig_vfd_its,fig_st2_its,
  labels = c("A", "B", "C"),
  label_size = 14,
  label_fontface = "bold",
  ncol = 3,
  rel_widths = c(1, 1, 1)
)

print(fig_its_combined_lunginjury)

# Save as PDF
ggsave(
  file.path(output_dir, "figures", "Figure_ITS_Biomarkers_Combined.pdf"),
  plot = fig_its_combined_lunginjury,
  width = 12,
  height = 4.5,
  dpi = 300
)



```


```{r candida_its_by_culture_figures}

library(tidyverse)
library(ggpubr)
library(ggbeeswarm)

# Merge ITS data with culture results
factsalir_cxpos_its_rel <- merge(factsalir_cxpos, relative_3cat, by = "SubjectID", all.x = TRUE)
factsalir_cxpos_its_abs <- merge(factsalir_cxpos, absolute_3cat, by = "SubjectID", all.x = TRUE)

# Define color palette and comparisons
culture_colors <- c("BacterialPathogen" = "#00A087", "NRF" = "#3C5488", "Candida" = "#E64B35")
culture_labels <- c("BacterialPathogen" = "Bacterial\nPathogen", 
                    "NRF" = "Normal\nFlora", 
                    "Candida" = "Candida")

pairwise_comparisons <- list(c("Candida", "NRF"), c("Candida", "BacterialPathogen"))

# ===== FIGURE 1: C. albicans Relative Abundance by Culture =====
fig_calb_rel <- factsalir_cxpos_its_rel %>%
  filter(!is.na(Calbicans_pct), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = Calbicans_pct, fill = cxres_lrt_f, color = cxres_lrt_f)) +
  
  geom_violin(alpha = 0.6, size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", 
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  
  labs(
    title = "C. albicans Relative Abundance by Culture Results",
    x = "",
    y = "C. albicans Relative Abundance (%)"
  ) +
  theme_pubr() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 13, face = "bold")
  )

print(fig_calb_rel)

ggsave(
  file.path(output_dir, "figures", "Figure_Calbicans_RelAbund_by_Culture.pdf"),
  plot = fig_calb_rel,
  width = 6,
  height = 5.5,
  dpi = 300
)


# ===== FIGURE 2: C. albicans Absolute Reads by Culture =====
fig_calb_abs <- factsalir_cxpos_its_abs %>%
  filter(!is.na(Calbicans_reads), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = (Calbicans_reads+1), fill = cxres_lrt_f, color = cxres_lrt_f)) +
  
  geom_violin(alpha = 0.6, size = 0.5) +
  geom_beeswarm(aes(color = cxres_lrt_f), alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  
  stat_compare_means(
    comparisons = pairwise_comparisons,
    method = "wilcox.test",
    label = "p.format", 
    p.format = list(accuracy = 0.001, scale = 1),
    tip.length = 0.02,
    step.increase = 0.1,
    size = 3.5,
    bracket.size = 0.4
  ) +
  
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  
  labs(
    title = "C. albicans Absolute Abundance by Culture Results",
    x = "",
    y = "C. albicans ITS Reads (log10)"
  ) +
  theme_pubr() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 13, face = "bold")
  )

print(fig_calb_abs)

ggsave(
  file.path(output_dir, "figures", "Figure_Calbicans_AbsReads_by_Culture.pdf"),
  plot = fig_calb_abs,
  width = 6,
  height = 5.5,
  dpi = 300
)

```


################################################################################
# qPCR DATA INTEGRATION
#
# Purpose:
#   1. Load qPCR dataset
#   2. Merge with facts_its_rel_analysis (ITS cohort)
#   3. Create correlation: qPCR vs ITS relative abundance
#
################################################################################

```{r qpcr_load and corr plot}
qpcr_baseline_eta <-read_csv("qpcr_baseline_eta_6.27.26.csv")

# Merge with facts_its_rel_analysis
facts_its_qpcr <- facts_its_abs_analysis %>%
  left_join(qpcr_baseline_eta, by = "SubjectIDDay")

facts_its_qpcr$qPCR_fungal_conc <-as.numeric(facts_its_qpcr$qPCR_fungal_conc)

# Publication-quality plot with Spearman correlation
correlation_plot_final <- facts_its_qpcr %>%
 # filter(Total_Reads > 0 & qPCR_fungal_conc > 0) %>%  # Remove zeros for log scale
  ggplot(aes(x = Total_Reads+1, y = qPCR_fungal_conc+1)) + 
  geom_point(alpha = 0.6, size = 2.5, color = "#2166AC") +
  geom_smooth(method = "loess", color = "#B2182B", fill = "#B2182B", 
              alpha = 0.2, linewidth = 1, span = 0.75) + stat_cor() + scale_y_log10() + scale_x_log10()+
  labs(
    x = "Total Fungal ITS Reads (log10)",
    y = "18S rRNA Gene Copies/μL (qPCR, log10)",
    title = "Correlation Between ITS Sequencing and qPCR Fungal Quantification"
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
    axis.title = element_text(face = "bold", size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
    legend.position = "none"
  ) +
  annotation_logticks(sides = "bl", size = 0.3)

print(correlation_plot_final)
ggsave("correlation_plot_final.pdf", width=8, height = 6)


```


```{r uw_validation_three_panel_IMPROVED}

library(tidyverse)
library(ggpubr)
library(ggbeeswarm)
library(patchwork)

# Load UW data
uwbal_clean <- read_csv("uwbal_clean_UsedForFacts_9.30.25.csv")

# Rename Yeast to Candida and set order
uwbal_clean <- uwbal_clean %>%
  mutate(all_bugs_grouped = as.character(all_bugs_grouped)) %>%
  mutate(all_bugs_grouped = if_else(all_bugs_grouped == "Yeast", "Candida", all_bugs_grouped)) %>%
  mutate(all_bugs_grouped = ordered(all_bugs_grouped, 
                                     levels = c("Saureus", "NoGrowth", "Candida")))

# Filter out NRF (only 3 cases)
uwbal_clean <- uwbal_clean %>% filter(all_bugs_grouped != "NRF")

# Define comparisons
my_comparisons <- list(c("Candida", "NoGrowth"), c("Candida", "Saureus"))

# Color palette matching UPMC culture results
uw_colors <- c("Candida" = "#E64B35", "Saureus" = "#00A087", "NoGrowth" = "#3C5488")

# PANEL A: Driving Pressure
fig_a <- uwbal_clean %>%
  filter(bronch_drivingpressure_8am_calculated_2 > 0) %>%
  ggplot(aes(x = all_bugs_grouped, y = bronch_drivingpressure_8am_calculated_2, 
             fill = all_bugs_grouped, color = all_bugs_grouped)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = uw_colors) + scale_color_manual(values = uw_colors) +
  labs(title = "Driving Pressure", x = "", y = "Driving Pressure (cmH2O)") +
  theme_pubr() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"))

# PANEL B: VFD
fig_b <- uwbal_clean %>%
  ggplot(aes(x = all_bugs_grouped, y = vfds_bronch, 
             fill = all_bugs_grouped, color = all_bugs_grouped)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = uw_colors) + scale_color_manual(values = uw_colors) +
  labs(title = "Ventilator-Free Days", x = "", y = "VFD (days)") +
  theme_pubr() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"))

# PANEL C: ICU Mortality
mort_pct <- uwbal_clean %>%
  group_by(all_bugs_grouped) %>%
  summarise(deaths = sum(icu_mortality == 1, na.rm=TRUE),
            total = n(),
            pct_mortality = deaths / total * 100, .groups = 'drop')

comparisons <- list(c("Candida", "Saureus"), c("Candida", "NoGrowth"))
pvals <- sapply(comparisons, function(pair) {
  mat <- table(uwbal_clean$icu_mortality, uwbal_clean$all_bugs_grouped)[, pair]
  fisher.test(mat)$p.value
})

comparison_results <- data.frame(
  group1 = sapply(comparisons, `[`, 1),
  group2 = sapply(comparisons, `[`, 2),
  p_value = pvals,
  p_label = signif(pvals, 3)
)

fig_c <- ggplot(mort_pct, aes(x = all_bugs_grouped, y = pct_mortality, fill = all_bugs_grouped)) +
  geom_col(alpha = 0.7, color = "black", size = 0.5) +
  geom_text(aes(label = paste0(round(pct_mortality), "%")), vjust = -0.5, fontface = "bold", size = 4) +
  scale_fill_manual(values = uw_colors) +
  labs(title = "ICU Mortality", x = "", y = "ICU Mortality (%)") +
  ylim(0, max(mort_pct$pct_mortality) + 20) +
  annotate("text", x = 1, y = max(mort_pct$pct_mortality) + 5, 
           label = paste0("p = ", comparison_results$p_label[1]), size = 3.5) +
  annotate("text", x = 2, y = max(mort_pct$pct_mortality) + 10, 
           label = paste0("p = ", comparison_results$p_label[2]), size = 3.5) +
  theme_pubr() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"))

# Combine with patchwork
fig_uw <- fig_a + fig_b + fig_c +
  plot_annotation(tag_levels = list(c('A', 'B', 'C')),
                  title = "Respiratory Physiology and Outcome by BAL Culture Results in the UW Cohort",
                  theme = theme(plot.title = element_text(size = 20, face = "bold")))

print(fig_uw)

ggsave(file.path(output_dir, "figures", "Figure_UW_Validation.pdf"),
       plot = fig_uw, width = 18, height = 5, dpi = 300)



# IL-8 variable
uwbal_clean <- uwbal_clean %>%
  mutate(il8_bal = `balf_IL-8_proinf_V1_imputed`)

# Color palette matching UPMC culture results
uw_colors <- c("Candida" = "#E64B35", "Saureus" = "#00A087", "NoGrowth" = "#3C5488")

# Comparisons
my_comparisons <- list(c("Candida", "NoGrowth"), c("Candida", "Saureus"))

# IL-8 Figure
fig_il8_uw <- uwbal_clean %>%
  filter(!is.na(il8_bal)) %>%
  ggplot(aes(x = all_bugs_grouped, y = il8_bal, 
             fill = all_bugs_grouped, color = all_bugs_grouped)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = uw_colors) + scale_color_manual(values = uw_colors) +
  scale_y_log10() +
  labs(title = "BAL IL-8 by BAL Culture Results in UW Cohort", 
       x = "", y = "IL-8 (pg/mL, log10)") +
  theme_pubr() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"))

print(fig_il8_uw)

ggsave(file.path(output_dir, "figures", "Figure_UW_IL8.pdf"),
       plot = fig_il8_uw, width = 6, height = 6, dpi = 300)
```


```{r eta_neutrophil_elastase_updated}

library(tidyverse)
library(ggpubr)
library(ggbeeswarm)

# Define color palette and comparisons (from UPMC culture analysis)
culture_colors <- c("BacterialPathogen" = "#00A087", "NRF" = "#3C5488", "Candida" = "#E64B35")
culture_labels <- c("BacterialPathogen" = "Bacterial\nPathogen", 
                    "NRF" = "Normal\nFlora", 
                    "Candida" = "Candida")

my_comparisons <- list(c("Candida", "NRF"), c("Candida", "BacterialPathogen"))

# Neutrophil Elastase Figure
fig_eta_neutelas <- factsalir_cxpos %>%
  filter(!is.na(NeutElas_eta), !is.na(totProt_ETA), !is.na(cxres_lrt_f)) %>%
  ggplot(aes(x = cxres_lrt_f, y = NeutElas_eta / totProt_ETA, 
             fill = cxres_lrt_f, color = cxres_lrt_f)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = culture_colors, labels = culture_labels) +
  scale_color_manual(values = culture_colors, labels = culture_labels) +
  scale_y_log10() +
  labs(title = "LRT Neutrophil Elastase by Culture Results",
       x = "", y = "Neutrophil Elastase / Total Protein (log10)") +
  theme_pubr() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"))

print(fig_eta_neutelas)

ggsave(file.path(output_dir, "figures", "Figure_ETA_NeutrophilElastase_by_Culture.pdf"),
       plot = fig_eta_neutelas, width = 6, height = 6, dpi = 300)


```



```{r its_biomarker_il8_ang2_tnfr1_combined}

library(tidyverse)
library(ggpubr)
library(ggbeeswarm)
library(patchwork)

# Color palette and labels for C. albicans dominance
its_colors <- c("C.albicans_Dominance" = "brown3",
                "Candida_Other_Dominance" = "pink2",
                "No_Candida_Dominance" = "gray70")

its_labels <- c("C.albicans_Dominance" = "C. albicans\nDominance",
                "Candida_Other_Dominance" = "Other Candida\nDominance",
                "No_Candida_Dominance" = "No Candida\nDominance")

# Comparisons
my_comparisons <- list(
  c("C.albicans_Dominance", "Candida_Other_Dominance"),
  c("C.albicans_Dominance", "No_Candida_Dominance")
)

# PANEL A: IL-8
fig_a <- facts_its_rel_analysis %>%
  filter(!is.na(il8), !is.na(Candida_Dominance_3cat)) %>%
  ggplot(aes(x = Candida_Dominance_3cat, y = il8,
             fill = Candida_Dominance_3cat, color = Candida_Dominance_3cat)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = its_colors, labels = its_labels) +
  scale_color_manual(values = its_colors, labels = its_labels) +
  scale_x_discrete(labels = its_labels) +
  scale_y_log10() +
  labs(title = "plasma IL-8", x = "", y = "IL-8 (pg/mL, log10)") +
  theme_pubr() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 9, face = "bold"),
        plot.title = element_text(size = 11, face = "bold"))

# PANEL B: Angiopoietin-2
fig_b <- facts_its_rel_analysis %>%
  filter(!is.na(ang2), !is.na(Candida_Dominance_3cat)) %>%
  ggplot(aes(x = Candida_Dominance_3cat, y = ang2,
             fill = Candida_Dominance_3cat, color = Candida_Dominance_3cat)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = its_colors, labels = its_labels) +
  scale_color_manual(values = its_colors, labels = its_labels) +
  scale_x_discrete(labels = its_labels) +
  scale_y_log10() +
  labs(title = "plasma Angiopoietin-2", x = "", y = "Ang-2 (pg/mL, log10)") +
  theme_pubr() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 9, face = "bold"),
        plot.title = element_text(size = 11, face = "bold"))

# PANEL C: TNFR1
fig_c <- facts_its_rel_analysis %>%
  filter(!is.na(tnfr1_final), !is.na(Candida_Dominance_3cat)) %>%
  ggplot(aes(x = Candida_Dominance_3cat, y = tnfr1_final,
             fill = Candida_Dominance_3cat, color = Candida_Dominance_3cat)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = its_colors, labels = its_labels) +
  scale_color_manual(values = its_colors, labels = its_labels) +
  scale_x_discrete(labels = its_labels) +
  scale_y_log10() +
  labs(title = "plasma sTNFR1", x = "", y = "sTNFR1 (pg/mL, log10)") +
  theme_pubr() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 9, face = "bold"),
        plot.title = element_text(size = 11, face = "bold"))

# Combine into single row
fig_its_inflammatory <- fig_a + fig_b + fig_c +
  plot_annotation(tag_levels = list(c('A', 'B', 'C')),
                  title = "Inflammatory Biomarkers by C. albicans Dominance",
                  theme = theme(plot.title = element_text(size = 14, face = "bold")))

print(fig_its_inflammatory)

ggsave(file.path(output_dir, "figures", "Figure_ITS_Inflammatory_Biomarkers_Combined.pdf"),
       plot = fig_its_inflammatory, width = 12, height = 4.5, dpi = 300)


```


```{r clusters_updated_publication}

library(tidyverse)
library(ggpubr)
library(ggbeeswarm)
library(survival)
library(survminer)
library(patchwork)

# Load and prepare cluster data
Mbiome_SampleMetadata_2.24.23 <- read_csv("Mbiome_SampleMetadata_2.24.23.csv")
MultiCompMbio_all_clusters_allCohorts_2 <- read_csv("MultiCompMbio_all_clusters_allCohorts_2.csv")

clustersMbio <- merge(Mbiome_SampleMetadata_2.24.23, MultiCompMbio_all_clusters_allCohorts_2, 
                      by = "sample_id", all = TRUE)

clustersMbio_lung_its <- clustersMbio %>%
  filter(SampleType.x == "ETA") %>%
  filter(SubjectID %in% facts_its_rel_analysis$SubjectID) %>%
  filter(Day == 1) %>%
  group_by(SubjectID) %>%
  slice_max(total16sreads, n = 1, with_ties = FALSE) %>%
  ungroup()

# Merge with facts_its
facts_its_clusters <- merge(facts_its_rel_analysis, clustersMbio_lung_its, by = "SubjectID", all.x = TRUE)

# Define comparisons
my_comparisons <- list(c("1", "2"), c("1", "3"))

# Color palette for clusters
cluster_colors <- c("1" = "brown3", "2" = "pink2", "3" = "steelblue")

# ===== PANEL A: Survival Curve =====
survival_fit <- survfit(Surv(survival_time90_fromICUAdmission, survival_status90) ~ 
                        lung_bacterial_fungal_cluster, 
                        data = facts_its_clusters)

fig_surv <- ggsurvplot(survival_fit, 
                       pval = TRUE,
                       pval.method = TRUE,
                       title = "90-Day Survival by Clusters",
                       legend.labs = c("Cluster 1", "Cluster 2", "Cluster 3"),
                       palette = c("brown3", "pink2", "steelblue"),
                       risk.table = TRUE,
                       tables.theme = clean_theme(),
                       legend.title = "Clusters",
                       font.legend = c(10, "bold", "black"),
                       font.title = c(12, "bold", "black"),
                       xlab = "Days Post ICU Admission",
                       ylab = "Probability of Survival",   
                       xlim = c(0, 90),  # ← LIMIT X-AXIS TO 90 DAYS
                       break.time.by = 15,  # ← NICE SPACING
                       size = 1.2)

# ===== PANEL B: C. albicans Abundance =====
fig_calb <- facts_its_clusters %>%
  filter(!is.na(lung_bacterial_fungal_cluster)) %>%
  ggplot(aes(x = as.factor(lung_bacterial_fungal_cluster), y = Calbicans_pct,
             fill = as.factor(lung_bacterial_fungal_cluster),
             color = as.factor(lung_bacterial_fungal_cluster))) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  labs(title = "C. albicans Abundance",
       x = "Bacterial-Fungal Cluster", y = "C. albicans Relative Abundance (%)") +
  theme_pubr() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 11, face = "bold"),
        plot.title = element_text(size = 12, face = "bold"))

# ===== PANEL C: Bacterial Pathogen Abundance =====
fig_pathogen <- facts_its_clusters %>%
  filter(!is.na(lung_bacterial_fungal_cluster)) %>%
  ggplot(aes(x = as.factor(lung_bacterial_fungal_cluster), y = PathogenSum,
             fill = as.factor(lung_bacterial_fungal_cluster),
             color = as.factor(lung_bacterial_fungal_cluster))) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = NA, size = 0.5) +
  geom_beeswarm(alpha = 0.5, size = 2, cex = 2, corral = "wrap") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",
                     label = "p.format", p.format = list(accuracy = 0.001, scale = 1),
                     tip.length = 0.02, step.increase = 0.1, size = 3.5, bracket.size = 0.4) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  labs(title = "Bacterial Pathogen Abundance",
       x = "Bacterial-Fungal Cluster", y = "Pathogen Relative Abundance (%)") +
  theme_pubr() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 11, face = "bold"),
        plot.title = element_text(size = 12, face = "bold"))

# ===== COMBINE: Abundance panels (top) + Survival (bottom) =====
abundance_combined <- fig_calb + fig_pathogen +
  plot_annotation(tag_levels = list(c('A', 'B')),
                  theme = theme(plot.title = element_text(size = 13, face = "bold")))

# Survival with risk table
surv_with_risk <- plot_grid(
  fig_surv$plot,
  fig_surv$table,
  ncol = 1,
  align = "v",
  axis = "lr",
  rel_heights = c(3, 1)
)

# Final combined figure
fig_clusters_final <- (abundance_combined / surv_with_risk) +
  plot_annotation(title = "Integrated Bacterial-Fungal Clusters and Clinical Outcomes",
                  theme = theme(plot.title = element_text(size = 14, face = "bold")))

print(fig_clusters_final)

# Save
ggsave(file.path(output_dir, "figures", "Figure_Clusters_Combined.pdf"),
       plot = fig_clusters_final, width = 10, height = 10, dpi = 300)

```

```{r clusters and cultures}
# ===== LOAD AND PREPARE DATA =====
clustersMbio_lung_cxpos <- clustersMbio %>%
  filter(SampleType.x == "ETA") %>%
  filter(SubjectID %in% factsalir_cxpos$SubjectID) %>%
  filter(Day == 1) %>%
  group_by(SubjectID) %>%
  slice_max(total16sreads, n = 1, with_ties = FALSE) %>%
  ungroup()

clustersMbio_lung_cxpos <- merge(clustersMbio_lung_cxpos, factsalir_cxpos, 
                                 by = "SubjectID", all.x = TRUE)


# ===== CONTINGENCY TABLE =====
culture_cluster_table <- table(clustersMbio_lung_cxpos$cxres_lrt_f, 
                              clustersMbio_lung_cxpos$lung_bacterial_fungal_cluster)

cat("\n════════════════════════════════════════════════════════════════════════════\n")
cat("CULTURE RESULTS vs. BACTERIAL-FUNGAL CLUSTERS\n")
cat("════════════════════════════════════════════════════════════════════════════\n\n")
print(culture_cluster_table)

# Calculate percentages
culture_cluster_pct <- prop.table(culture_cluster_table, margin = 1) * 100

cat("\n\nPERCENTAGES WITHIN EACH CULTURE CATEGORY:\n\n")
print(round(culture_cluster_pct, 1))

# ===== PREPARE DATA FOR VISUALIZATION =====
culture_cluster_data <- as.data.frame(culture_cluster_table) %>%
  rename(Culture = Var1, Cluster = Var2, Count = Freq) %>%
  group_by(Culture) %>%
  mutate(Pct = 100 * Count / sum(Count),
         Total = sum(Count)) %>%
  ungroup()

# ===== STACKED BAR CHART =====
fig_culture_cluster <- culture_cluster_data %>%
  mutate(Culture = factor(Culture, levels = c("NRF", "BacterialPathogen", "Candida")),
         Cluster = factor(Cluster, levels = c("1", "2", "3"))) %>%
  ggplot(aes(x = Culture, y = Count, fill = Cluster)) +
  geom_col(alpha = 0.8, color = "black", size = 0.5) +
  
  # Add percentage labels
  geom_text(aes(label = sprintf("%.0f%%\n(n=%d)", Pct, Count)),
            position = position_stack(vjust = 0.5),
            fontface = "bold", size = 3.5, color = "white") +
  
  scale_fill_manual(
    values = c("1" = "brown3", "2" = "pink2", "3" = "steelblue")
  ) +
  
  scale_y_continuous(expand = c(0, 0)) +
  
  labs(
    title = "Clinical Culture Categories and Integrative Bacterial-Fungal Clusters",
        x = "Clinical Culture Result",
    y = "Number of Subjects",
    fill = "16S-ITS Clusters"
  ) +
  
  theme_pubr() +
  theme(
    plot.title = element_text(size = 13, face = "bold"),
    plot.subtitle = element_text(size = 11, face = "italic"),
    axis.title = element_text(size = 11, face = "bold"),
    legend.title = element_text(size = 10, face = "bold"),
    legend.position = "right"
  )

print(fig_culture_cluster)

# ===== SAVE =====
ggsave(
  file.path(output_dir, "figures", "FigureSUPP_Culture_vs_Clusters_Overlap.pdf"),
  plot = fig_culture_cluster,
  width = 8,
  height = 5.5,
  dpi = 300
)


```
