---
title: "FACTS: Murine In Vivo and In Vitro Analyses"
subtitle: "C. albicans Lung Injury Models — Publication Analysis Code"
author: "Kitsios Lab, University of Pittsburgh"
date: "July 2026"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: cosmo
---

# FACTS In Vivo / In Vitro Analysis

Analysis code for murine intratracheal *C. albicans* models and in vitro lung epithelial assays.

**Required input files:**

- `FACTS_rawdata_allExp_06.29.26.xlsx` — Master experimental dataset
- `FACTS_histologyScores_12.5.24_merged.xlsx` — Two-hit histology scoring
- `FACTS_histologyScores_4.9.25_LS.xlsx` — Single-hit histology scoring
- `humanInvitroData.xlsx` — Human in vitro cytotoxicity and TEER data

---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)
```

```{r load_libraries}
# =============================================================================
# Required packages
# =============================================================================
library(dplyr)
library(readxl)
library(readr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
library(ggpubr)
library(ggsci)
library(patchwork)
library(scales)
library(survival)
library(survminer)
library(nlme)
```

## Data Import and Processing

```{r load_data}
# =============================================================================
# Data Import and Derived Variables
# =============================================================================

# --- Load raw data ---
facts <- read_excel("FACTS_rawdata_allExp_06.29.26.xlsx")
histology2hit <- read_excel("FACTS_histologyScores_12.5.24_merged.xlsx")
histology1hit <- read_excel("FACTS_histologyScores_4.9.25_LS.xlsx")

# --- Normalized lung weight (left lung / body weight on sacrifice day) ---
facts <- facts %>%
  mutate(LungWeightNormalized = case_when(
    Sacrifice_day == 1 ~ LeftLung_weight / bodyweightD1,
    Sacrifice_day == 2 ~ LeftLung_weight / bodyweightD2,
    Sacrifice_day == 3 ~ LeftLung_weight / bodyweightD3,
    Sacrifice_day == 4 ~ LeftLung_weight / bodyweightD4,
    Sacrifice_day == 7 ~ LeftLung_weight / bodyweightD7
  ))

# --- Weight loss fraction per day (relative to Day 0) ---
for (d in 1:7) {
  wt_col <- paste0("bodyweightD", d)
  facts[[paste0("d", d, "weightloss")]] <-
    (facts$bodyweightD0 - facts[[wt_col]]) / facts$bodyweightD0
}

# --- Percent body weight per day (1 = baseline) ---
facts$d0weightpercent <- 1
for (d in 1:7) {
  facts[[paste0("d", d, "weightpercent")]] <-
    1 - facts[[paste0("d", d, "weightloss")]]
}

# --- Total BAL neutrophils ---
facts$BAL_Neutrophils_total <-
  (facts$BALNeutrophils_percent * facts$BALcellcount) / 100

# --- Experiment summary ---
experiment_summary <- facts %>%
  group_by(experiment) %>%
  summarise(
    n = n(),
    groups = paste(sort(unique(group)), collapse = " | "),
    genotypes = paste(sort(unique(genotype)), collapse = " | "),
    sex = paste(sort(unique(sex)), collapse = " | "),
    sacrifice_days = paste(sort(unique(Sacrifice_day)), collapse = " | "),
    .groups = "drop"
  )

print(experiment_summary, n = 50)
```

---

## Figure 2: Intratracheal *C. albicans* Causes Acute Lung Injury

```{r figure2}
# =============================================================================
# Figure 2: CAN6 vs Vehicle (Day 1 & 2)
# Experiments 2, 9, 10, 11, 14, 21 — WT mice
# =============================================================================

# --- Data subset ---
fig2_data <- facts %>%
  filter(experiment %in% c(2, 9, 10, 11, 14, 21),
         group %in% c("CAN6", "Veh"),
         genotype == "WT") %>%
  mutate(group = factor(group, levels = c("Veh", "CAN6")))

# --- Shared aesthetics ---
fig2_colors <- c("Veh" = "gray40", "CAN6" = "#D62728")
fig2_comparisons <- list(c("Veh", "CAN6"))
sacrifice_day_labels <- c("1" = "Day 1", "2" = "Day 2")

# --- Common theme ---
theme_fig2 <- function() {
  theme_pubr(base_size = 10) %+replace%
    theme(
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(size = 10, face = "bold"),
      axis.title.y = element_text(size = 9, angle = 90),
      axis.text.x = element_text(size = 9),
      axis.text.y = element_text(size = 8),
      plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
      plot.margin = margin(4, 4, 4, 4)
    )
}

# --- Helper: faceted boxplot + beeswarm ---
fig2_boxplot <- function(data, yvar, ylab, title,
                         ylimits = NULL, ybreaks = waiver()) {
  p <- ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = fig2_comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = fig2_colors) +
    facet_grid(~ Sacrifice_day,
               labeller = labeller(Sacrifice_day = sacrifice_day_labels)) +
    labs(x = NULL, y = ylab, title = title) +
    theme_fig2()
  if (!is.null(ylimits)) {
    p <- p + scale_y_continuous(limits = ylimits, breaks = ybreaks)
  }
  p
}

# --- Panel B: Weight trajectory ---
fig2_wt_long <- fig2_data %>%
  dplyr::select(mouseID, group, d0weightpercent, d1weightpercent, d2weightpercent) %>%
  pivot_longer(cols = d0weightpercent:d2weightpercent,
               names_to = "Day", values_to = "WeightPct") %>%
  mutate(Day = factor(Day,
    levels = c("d0weightpercent", "d1weightpercent", "d2weightpercent"),
    labels = c("Day 0", "Day 1", "Day 2")))

fig2_wt_summary <- fig2_wt_long %>%
  group_by(group, Day) %>%
  summarise(
    median = median(WeightPct, na.rm = TRUE),
    q25 = quantile(WeightPct, 0.25, na.rm = TRUE),
    q75 = quantile(WeightPct, 0.75, na.rm = TRUE),
    .groups = "drop")

fig2B <- ggplot(fig2_wt_summary,
                aes(x = Day, y = median, group = group, color = group)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.15, linewidth = 0.5) +
  scale_color_manual(values = fig2_colors) +
  scale_y_continuous(limits = c(0.9, 1.05), labels = scales::percent) +
  labs(x = NULL, y = "Body Weight (% baseline)", color = NULL) +
  theme_pubr(base_size = 10) +
  theme(legend.position = "top", legend.key.size = unit(0.4, "cm"),
        plot.margin = margin(4, 4, 4, 4),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.3)) +
  ggtitle("Weight Change by Group")

# --- Panels C–I: Faceted boxplots ---
fig2_data$BAL_IgM <- as.numeric(fig2_data$BAL_IgM)

fig2C <- fig2_boxplot(fig2_data, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight", ylimits = c(0, 7))
fig2D <- fig2_boxplot(fig2_data, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein")
fig2E <- fig2_boxplot(fig2_data, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM",
  ylimits = c(0, 8000), ybreaks = seq(0, 8000, 2000))
fig2F <- fig2_boxplot(fig2_data, "BAL_RAGE",
  ylab = "BAL sRAGE (pg/ml)", title = "BAL sRAGE",
  ylimits = c(0, 20000), ybreaks = seq(0, 20000, 5000))
fig2G <- fig2_boxplot(fig2_data, "BALcellcount",
  ylab = expression("BAL Cells (" * 10^4 * ")"), title = "BAL Cell Count",
  ylimits = c(0, 800), ybreaks = seq(0, 800, 200))
fig2H <- fig2_boxplot(fig2_data, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils",
  ylimits = c(0, 800), ybreaks = seq(0, 800, 200))
fig2I <- fig2_boxplot(fig2_data, "NeutElastase",
  ylab = "BAL NE (pg/ml)", title = "Neutrophil Elastase",
  ylimits = c(0, 12000), ybreaks = seq(0, 12000, 4000))

# --- Panel J: Histology score ---
histology1hit$group <- case_when(
  histology1hit$Mouse %in% c("m1","m3","m7","m8") ~ "Veh",
  histology1hit$Mouse %in% c("m2","m4","m5","m6","m9") ~ "CAN6",
  TRUE ~ NA_character_
)
histology1hit <- histology1hit %>% filter(!is.na(group))
histology1hit$group <- factor(histology1hit$group, levels = c("Veh", "CAN6"))

fig2J <- ggplot(histology1hit,
                aes(x = group, y = ScorePerField, color = group)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
  geom_jitter(width = 0.15, size = 1.2, alpha = 0.6) +
  stat_compare_means(comparisons = fig2_comparisons, size = 3,
                     label = "p.format", bracket.size = 0.4) +
  scale_color_manual(values = fig2_colors) +
  labs(x = NULL, y = "Histology Score", title = "Histology Score") +
  theme_fig2()

# --- Placeholders ---
fig2A <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "(schematic)", size = 5, color = "grey60")
fig2K <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "(histology images)", size = 5, color = "grey60")

# --- Assembly ---
fig2_top    <- fig2A + fig2B + fig2C + plot_layout(nrow = 1, widths = c(1.5, 1.2, 0.8))
fig2_middle <- fig2D + fig2E + fig2F + fig2I + plot_layout(nrow = 1)
fig2_bottom <- fig2G + fig2H + fig2J + fig2K +
  plot_layout(nrow = 1, widths = c(1, 1, 0.7, 1.3))

fig2 <- fig2_top / fig2_middle / fig2_bottom +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("Figure2_R.pdf", fig2, width = 10, height = 10)
```

## Supplemental: CFUs and Histology Features (CAN6 vs Veh)

```{r supp_cfus_histology}
# =============================================================================
# Supplemental: C. albicans CFUs (Day 2, CAN6 vs Veh)
# =============================================================================

fig2_cfu_data <- fig2_data %>%
  filter(Sacrifice_day == 2) %>%
  dplyr::select(mouseID, group, LeftLung_CanCFUs, BAL_CanCFUs) %>%
  pivot_longer(cols = c(LeftLung_CanCFUs, BAL_CanCFUs),
               names_to = "Compartment", values_to = "CFUs") %>%
  mutate(
    log10CFU = log10(CFUs + 1),
    Compartment = factor(Compartment,
      levels = c("BAL_CanCFUs", "LeftLung_CanCFUs"),
      labels = c("BAL (CFUs/ml)", "Lung (CFUs/g)"))
  )

fig_supp_CFUs <- ggplot(fig2_cfu_data,
    aes(x = group, y = log10CFU, color = group)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
  geom_beeswarm(size = 1.2, cex = 2) +
  stat_compare_means(comparisons = fig2_comparisons, size = 3,
                     label = "p.format", bracket.size = 0.4) +
  scale_color_manual(values = fig2_colors) +
  facet_wrap(~ Compartment) +
  labs(x = NULL,
       y = expression(log[10] ~ italic("C. albicans") ~ "CFUs"),
       title = expression(italic("C. albicans") ~ "Recovery (Day 2)")) +
  theme_fig2()

ggsave("FigS_CFUs_CAN6.pdf", fig_supp_CFUs, width = 5, height = 4)

# =============================================================================
# Supplemental: Histologic Injury Features (CAN6 vs Veh)
# =============================================================================

# Average across reviewers per field
histology1hit_avg <- histology1hit %>%
  group_by(Mouse, Field) %>%
  summarise(
    NeutrophilsAlveolarSpace = mean(NeutrophilsAlveolarSpace, na.rm = TRUE),
    NeutrophilsInterstitialSpace = mean(NeutrophilsInterstitialSpace, na.rm = TRUE),
    HyalineMembranes = mean(HyalineMembranes, na.rm = TRUE),
    ProteinaceousDebrisAirspaces = mean(ProteinaceousDebrisAirspaces, na.rm = TRUE),
    AlveolarSeptalThickening = mean(AlveolarSeptalThickening, na.rm = TRUE),
    ScorePerField = mean(ScorePerField, na.rm = TRUE),
    .groups = "drop"
  )

histology1hit_avg$group <- case_when(
  histology1hit_avg$Mouse %in% c("m1","m3","m7","m8") ~ "Veh",
  histology1hit_avg$Mouse %in% c("m2","m4","m5","m6","m9") ~ "CAN6",
  TRUE ~ NA_character_
)
histology1hit_avg <- histology1hit_avg %>% filter(!is.na(group))
histology1hit_avg$group <- factor(histology1hit_avg$group, levels = c("Veh", "CAN6"))

# --- Helper for histology panels ---
histo_panel <- function(data, yvar, ylab, title) {
  ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_jitter(width = 0.15, size = 1.2, alpha = 0.6) +
    stat_compare_means(comparisons = list(c("Veh", "CAN6")), size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = fig2_colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_pubr(base_size = 10) +
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          axis.text.x = element_text(size = 9),
          axis.text.y = element_text(size = 8),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
          plot.margin = margin(4, 4, 4, 4))
}

histA <- histo_panel(histology1hit_avg, "NeutrophilsAlveolarSpace",
  ylab = "Score", title = "Alveolar Neutrophils")
histB <- histo_panel(histology1hit_avg, "NeutrophilsInterstitialSpace",
  ylab = "Score", title = "Interstitial Neutrophils")
histC <- histo_panel(histology1hit_avg, "HyalineMembranes",
  ylab = "Score", title = "Hyaline Membranes")
histD <- histo_panel(histology1hit_avg, "ProteinaceousDebrisAirspaces",
  ylab = "Score", title = "Proteinaceous Debris")
histE <- histo_panel(histology1hit_avg, "AlveolarSeptalThickening",
  ylab = "Score", title = "Alveolar Septal Thickening")

figS_histology <- histA + histB + histC + histD + histE +
  plot_layout(nrow = 1) +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("FigS_Histology_Features.pdf", figS_histology, width = 14, height = 4)
```

---

## Figure 3: Neutrophil Depletion

```{r figure3}
# =============================================================================
# Figure 3: Neutrophil depletion — PMN_DTR vs PMN_WT
# Experiments 8, 20, 23
# =============================================================================

fig3_data <- facts %>%
  filter(experiment %in% c(8, 20, 23)) %>%
  mutate(genotype = factor(genotype, levels = c("PMNWT", "PMNDTR")))

# Remove mice that did not receive inoculum or died after anesthesia
fig3_data <- fig3_data %>%
  filter(!(mouseID == "WTCAN6_m5" & experiment == 20),
         !(mouseID == "WTCAN6_m1" & experiment == 20),
         !(mouseID == "PMNDTR_m15" & experiment == 23),
         !(mouseID == "PMNDTR_m11" & experiment == 23))

# --- Shared aesthetics ---
fig3_colors <- c("PMNWT" = "royalblue", "PMNDTR" = "maroon")
fig3_comparisons <- list(c("PMNWT", "PMNDTR"))

theme_fig3 <- function() {
  theme_pubr(base_size = 10) %+replace%
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          axis.text.x = element_text(size = 9),
          axis.text.y = element_text(size = 8),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
          plot.margin = margin(4, 4, 4, 4))
}

fig3_boxplot <- function(data, yvar, ylab, title = NULL,
                         ylimits = NULL, ybreaks = waiver()) {
  p <- ggplot(data, aes(x = genotype, y = .data[[yvar]], color = genotype)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = fig3_comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = fig3_colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_fig3()
  if (!is.null(ylimits)) p <- p + scale_y_continuous(limits = ylimits, breaks = ybreaks)
  p
}

# --- Panel C: Weight trajectory ---
fig3_wt_long <- fig3_data %>%
  dplyr::select(mouseID, genotype, d0weightpercent, d1weightpercent, d2weightpercent) %>%
  pivot_longer(cols = d0weightpercent:d2weightpercent,
               names_to = "Day", values_to = "WeightPct") %>%
  mutate(Day = factor(Day,
    levels = c("d0weightpercent", "d1weightpercent", "d2weightpercent"),
    labels = c("Day 0", "Day 1", "Day 2")))

fig3_wt_summary <- fig3_wt_long %>%
  group_by(genotype, Day) %>%
  summarise(median = median(WeightPct, na.rm = TRUE),
            q25 = quantile(WeightPct, 0.25, na.rm = TRUE),
            q75 = quantile(WeightPct, 0.75, na.rm = TRUE), .groups = "drop")

fig3C <- ggplot(fig3_wt_summary,
                aes(x = Day, y = median, group = genotype, color = genotype)) +
  geom_line(linewidth = 0.8) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.15, linewidth = 0.5) +
  scale_color_manual(values = fig3_colors) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Body Weight (% baseline)", color = NULL) +
  theme_pubr(base_size = 10) +
  theme(legend.position = "top", legend.key.size = unit(0.4, "cm"),
        plot.margin = margin(4, 4, 4, 4),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.3))

# --- Endpoint data: remove deaths and survival-only experiment ---
fig3_endpoints <- fig3_data %>%
  filter(!(mouseID == "DTRCAN6_m2" & experiment == 20),
         !(mouseID %in% c("DTRCAN6_m7", "DTRCAN6_m8") & experiment == 8),
         !(mouseID %in% c("PMNDTR_m6", "PMNDTR_m10", "PMNDTR_m13", "PMNDTR_m18") & experiment == 23),
         experiment != 23)
fig3_endpoints$BAL_IgM <- as.numeric(fig3_endpoints$BAL_IgM)
fig3_endpoints <- fig3_endpoints %>%
  mutate(log10_lungCFU = log10(LeftLung_CanCFUs + 1),
         log10_kidneyCFU = log10(Rkidney_CanCFUs + 1))

# --- Panels D–J ---
fig3D <- fig3_boxplot(fig3_endpoints, "log10_lungCFU",
  ylab = expression(log[10] ~ "Lung" ~ italic("C. albicans") ~ "CFUs/g"),
  title = "Lung CFUs", ylimits = c(0, 7.5))
fig3E <- fig3_boxplot(fig3_endpoints, "log10_kidneyCFU",
  ylab = expression(log[10] ~ "Kidney" ~ italic("C. albicans") ~ "CFUs/g"),
  title = "Kidney CFUs", ylimits = c(0, 5))
fig3F <- fig3_boxplot(fig3_endpoints, "BALcellcount",
  ylab = expression("BAL Cells (" * 10^4 * ")"), title = "BAL Cell Count")
fig3G <- fig3_boxplot(fig3_endpoints, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils")
fig3H <- fig3_boxplot(fig3_endpoints, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight", ylimits = c(0, 10))
fig3I <- fig3_boxplot(fig3_endpoints, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein")
fig3J <- fig3_boxplot(fig3_endpoints, "BAL_RAGE",
  ylab = "BAL sRAGE (pg/ml)", title = "BAL sRAGE")

# --- Placeholders ---
fig3A <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "(schematic)", size = 5, color = "grey60")
fig3B <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "(survival curve)", size = 5, color = "grey60")

# --- Assembly ---
fig3_row1 <- fig3A + fig3B + plot_layout(widths = c(1.2, 1))
fig3_row2 <- fig3C + fig3D + fig3E + fig3F + plot_layout(nrow = 1)
fig3_row3 <- fig3G + fig3H + fig3I + fig3J + plot_layout(nrow = 1)

fig3 <- fig3_row1 / fig3_row2 / fig3_row3 +
  plot_layout(heights = c(1, 1.2, 1.2)) +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("Figure3_R.pdf", fig3, width = 10, height = 8.5)
```

---

## Figure 4: Live vs Heat-Killed *C. albicans*

```{r figure4}
# =============================================================================
# Figure 4: Live vs Heat-Killed — SC5314 + Clinical Isolate
# Row 1 (A–D): SC5314 (Experiments 30, 37, 38)
# Row 2 (E–H): Clinical Isolate (Experiment 40)
# =============================================================================

# --- SC5314 data ---
fig4_sc <- facts %>%
  filter(experiment %in% c(30, 37, 38), genotype == "WT") %>%
  mutate(group = case_when(
    group == "CAN6" ~ "Live CA",
    group %in% c("HKCAN6", "HKCAN") ~ "HKCA",
    TRUE ~ group)) %>%
  mutate(group = factor(group, levels = c("Live CA", "HKCA")))
fig4_sc$BAL_IgM <- as.numeric(fig4_sc$BAL_IgM)

# --- Clinical isolate data ---
fig4_clin <- facts %>%
  filter(experiment == 40, genotype == "WT") %>%
  mutate(group = case_when(
    group == "ClinCan6" ~ "Live CA",
    group == "HKClinCan6" ~ "HKCA",
    TRUE ~ group)) %>%
  mutate(group = factor(group, levels = c("Live CA", "HKCA")))
fig4_clin$BAL_IgM <- as.numeric(fig4_clin$BAL_IgM)

# --- Shared aesthetics ---
sc_colors   <- c("Live CA" = "darkorange2",  "HKCA" = "mediumpurple4")
clin_colors <- c("Live CA" = "firebrick",    "HKCA" = "steelblue3")
fig4_comparisons <- list(c("Live CA", "HKCA"))

theme_fig4 <- function() {
  theme_pubr(base_size = 10) %+replace%
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          axis.text.x = element_text(size = 9),
          axis.text.y = element_text(size = 8),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
          plot.margin = margin(4, 4, 4, 4))
}

fig4_boxplot <- function(data, yvar, ylab, title = NULL, colors,
                         ylimits = NULL, ybreaks = waiver()) {
  p <- ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = fig4_comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_fig4()
  if (!is.null(ylimits)) p <- p + scale_y_continuous(limits = ylimits, breaks = ybreaks)
  p
}

# --- Row 1: SC5314 ---
sc_A <- fig4_boxplot(fig4_sc, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils", colors = sc_colors)
sc_B <- fig4_boxplot(fig4_sc, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight", colors = sc_colors)
sc_C <- fig4_boxplot(fig4_sc, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein", colors = sc_colors)
sc_D <- fig4_boxplot(fig4_sc, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM", colors = sc_colors)

# --- Row 2: Clinical Isolate ---
clin_E <- fig4_boxplot(fig4_clin, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils", colors = clin_colors)
clin_F <- fig4_boxplot(fig4_clin, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight", colors = clin_colors)
clin_G <- fig4_boxplot(fig4_clin, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein", colors = clin_colors)
clin_H <- fig4_boxplot(fig4_clin, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM", colors = clin_colors)

# --- Title rows ---
title_sc <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "C. albicans SC5314",
           size = 4.5, fontface = "bold.italic")
title_clin <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "Clinical Isolate C. albicans",
           size = 4.5, fontface = "bold.italic")

# --- Assembly ---
row1 <- sc_A + sc_B + sc_C + sc_D + plot_layout(nrow = 1)
row2 <- clin_E + clin_F + clin_G + clin_H + plot_layout(nrow = 1)

fig4_combined <- wrap_elements(title_sc) / row1 / wrap_elements(title_clin) / row2 +
  plot_layout(heights = c(0.08, 1, 0.08, 1)) +
  plot_annotation(tag_levels = list(c("", "A","B","C","D", "", "E","F","G","H")),
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("Figure4_LiveVsHKCA.pdf", fig4_combined, width = 12, height = 7)
```

---

## Figure 5: Yeast-Lock + Human In Vitro

```{r figure5}
# =============================================================================
# Figure 5: Hyphal transition drives lung injury — Mouse + Human Data
# Row 1 (A–E): Mouse in vivo yeast-lock (Exps 13, 15)
# Row 2 (F–H): Human in vitro (LDH cytotoxicity, primary TEER, ALI TEER)
# =============================================================================

# --- Row 1: Mouse in vivo ---
fig5_data <- facts %>%
  filter(experiment %in% c(13, 15), genotype == "WT") %>%
  mutate(group = recode(group, "TT21" = "Wild-type", "TNRG1" = "Yeast-lock")) %>%
  mutate(group = factor(group, levels = c("Wild-type", "Yeast-lock")))

fig5_data <- fig5_data %>%
  filter(!(mouseID == "tt21_m3" & experiment == 13)) %>%
  filter(!(mouseID == "tt21_m2" & experiment == 15))
fig5_data$BAL_IgM <- as.numeric(fig5_data$BAL_IgM)
fig5_data <- fig5_data %>% mutate(log10_lungCFU = log10(LeftLung_CanCFUs + 1))

fig5_colors <- c("Wild-type" = "forestgreen", "Yeast-lock" = "goldenrod3")
fig5_comparisons <- list(c("Wild-type", "Yeast-lock"))

theme_fig5 <- function() {
  theme_pubr(base_size = 10) %+replace%
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          axis.text.x = element_text(size = 9),
          axis.text.y = element_text(size = 8),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
          plot.margin = margin(4, 4, 4, 4))
}

fig5_boxplot <- function(data, yvar, ylab, title = NULL,
                         ylimits = NULL, ybreaks = waiver()) {
  p <- ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = fig5_comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = fig5_colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_fig5()
  if (!is.null(ylimits)) p <- p + scale_y_continuous(limits = ylimits, breaks = ybreaks)
  p
}

fig5A <- fig5_boxplot(fig5_data, "log10_lungCFU",
  ylab = expression(log[10] ~ "Lung" ~ italic("C. albicans") ~ "CFUs/g"),
  title = "Lung CFUs", ylimits = c(0, 5))
fig5B <- fig5_boxplot(fig5_data, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils")
fig5C <- fig5_boxplot(fig5_data, "NeutElastase",
  ylab = "BAL NE (pg/ml)", title = "Neutrophil Elastase")
fig5D <- fig5_boxplot(fig5_data, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein")
fig5E <- fig5_boxplot(fig5_data, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM")

# --- Row 2: Human in vitro data ---
human_raw <- read_excel("humanInvitroData.xlsx", col_names = FALSE)

# Panel F: Human LDH Cytotoxicity by MOI
ldh_lysis <- as.numeric(human_raw[2, 2:17])
ldh_bg    <- as.numeric(human_raw[3, 2:17])
ldh_moi1  <- as.numeric(human_raw[4, 2:17])
ldh_moi5  <- as.numeric(human_raw[5, 2:17])
ldh_moi10 <- as.numeric(human_raw[6, 2:17])

ldh_df <- data.frame(
  value = c(ldh_lysis, ldh_bg, ldh_moi1, ldh_moi5, ldh_moi10),
  group = rep(c("Lysis_Control", "Background_Control", "MOI_1", "MOI_5", "MOI_10"), each = 16)
)
ldh_df$group <- factor(ldh_df$group,
  levels = c("Lysis_Control", "Background_Control", "MOI_1", "MOI_5", "MOI_10"))

fig5F <- ggplot(ldh_df, aes(x = group, y = value, color = group)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
  geom_beeswarm(size = 1.0, cex = 1.5) +
  stat_compare_means(comparisons = list(c("Lysis_Control", "MOI_1"),
                                        c("Lysis_Control", "MOI_5"),
                                        c("Lysis_Control", "MOI_10")),
                     size = 2.5, label = "p.signif", bracket.size = 0.3) +
  scale_color_manual(values = c("Lysis_Control" = "black", "Background_Control" = "orange",
                                "MOI_1" = "steelblue3", "MOI_5" = "mediumpurple",
                                "MOI_10" = "firebrick")) +
  scale_x_discrete(labels = c("Lysis", "Bkgd", "MOI 1", "MOI 5", "MOI 10")) +
  labs(x = NULL, y = "Relative Cytotoxicity", title = "Human Epithelial Cytotoxicity") +
  theme_fig5() + theme(axis.text.x = element_text(size = 7, angle = 30, hjust = 1))

# Panel G: Primary Human Bronchial TEER
teer1_veh <- as.numeric(human_raw[9, 2:13])
teer1_wt  <- as.numeric(human_raw[10, 2:13])
teer1_df <- data.frame(value = c(teer1_veh, teer1_wt),
                       group = rep(c("Vehicle", "Wild-type"), each = 12))
teer1_df$group <- factor(teer1_df$group, levels = c("Vehicle", "Wild-type"))

fig5G <- ggplot(teer1_df, aes(x = group, y = value, color = group)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
  geom_beeswarm(size = 1.2, cex = 2) +
  stat_compare_means(comparisons = list(c("Vehicle", "Wild-type")), size = 3,
                     label = "p.format", bracket.size = 0.4) +
  scale_color_manual(values = c("Vehicle" = "gray40", "Wild-type" = "forestgreen")) +
  labs(x = NULL, y = "Normalized TEER\n(relative to baseline)", title = "Primary Bronchial TEER") +
  theme_fig5()

# Panel H: Human Alveolar TEER (Vehicle vs WT vs Yeast-lock)
teer2_veh <- as.numeric(human_raw[13, 2:13])
teer2_wt  <- as.numeric(human_raw[14, 2:13])
teer2_yl  <- as.numeric(human_raw[15, 2:13])
teer2_df <- data.frame(value = c(teer2_veh, teer2_wt, teer2_yl),
                       group = rep(c("Vehicle", "Wild-type", "Yeast-lock"), each = 12))
teer2_df$group <- factor(teer2_df$group, levels = c("Vehicle", "Wild-type", "Yeast-lock"))

fig5H <- ggplot(teer2_df, aes(x = group, y = value, color = group)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
  geom_beeswarm(size = 1.2, cex = 2) +
  stat_compare_means(comparisons = list(c("Vehicle", "Wild-type"), c("Vehicle", "Yeast-lock")),
                     size = 3, label = "p.format", bracket.size = 0.4) +
  scale_color_manual(values = c("Vehicle" = "gray40", "Wild-type" = "forestgreen",
                                "Yeast-lock" = "goldenrod3")) +
  labs(x = NULL, y = "Normalized TEER\n(relative to baseline)", title = "Yeast-Lock Alveolar TEER") +
  theme_fig5()

# --- Assembly ---
row1 <- fig5A + fig5B + fig5C + fig5D + fig5E + plot_layout(nrow = 1)
row2 <- fig5F + fig5G + fig5H + plot_layout(nrow = 1, widths = c(1.3, 1, 1))

fig5 <- row1 / row2 +
  plot_layout(heights = c(1, 1)) +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("Figure5_combined.pdf", fig5, width = 14, height = 7)
```

---

## Supplemental: Candidalysin (Exp 31, Day 1)

```{r candidalysin_day1}
# =============================================================================
# Supplemental: Candidalysin mutant vs WT (BWP17 vs ece1)
# Experiment 31, Day 1 sacrifice
# =============================================================================

factsCL <- facts %>%
  filter(experiment == 31) %>%
  mutate(group = if_else(group == "BWP17", "WT", group),
         group = factor(group, levels = c("WT", "Ece1")))
factsCL$BAL_IgM <- as.numeric(factsCL$BAL_IgM)

cl_colors <- c("WT" = "gray28", "Ece1" = "#D62728")
cl_comparisons <- list(c("WT", "Ece1"))

# Weight trajectory
factsCL_wt <- factsCL %>%
  pivot_longer(cols = c(d0weightpercent, d1weightpercent),
               names_to = "Day", values_to = "WeightPct") %>%
  mutate(Day = factor(Day, levels = c("d0weightpercent", "d1weightpercent"),
                      labels = c("Day 0", "Day 1")))

cl_wt_sum <- factsCL_wt %>%
  group_by(group, Day) %>%
  summarise(median = median(WeightPct, na.rm = TRUE),
            q25 = quantile(WeightPct, 0.25, na.rm = TRUE),
            q75 = quantile(WeightPct, 0.75, na.rm = TRUE), .groups = "drop")

cl_wtplot <- ggplot(cl_wt_sum, aes(x = Day, y = median, group = group, color = group)) +
  geom_line(linewidth = 0.8) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.15, linewidth = 0.5) +
  scale_color_manual(values = cl_colors) +
  labs(x = NULL, y = "Body Weight (% baseline)", title = "Weight Change", color = NULL) +
  theme_pubr(base_size = 10) +
  theme(legend.position = "top", legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size = 10, face = "bold", hjust = 0.5))

# Endpoint panels
cl_boxplot <- function(data, yvar, ylab, title) {
  ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = cl_comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = cl_colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_pubr(base_size = 10) +
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5))
}

cl_neutrophils <- cl_boxplot(factsCL, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils")
cl_lungwt <- cl_boxplot(factsCL, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight")
cl_prot <- cl_boxplot(factsCL, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein")
cl_igm <- cl_boxplot(factsCL, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM")

figS_candidalysin <- cl_wtplot + cl_neutrophils + cl_lungwt + cl_prot + cl_igm +
  plot_layout(nrow = 1) +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("FigS_Candidalysin_Day1.pdf", figS_candidalysin, width = 14, height = 3.5)
```

---

## Figure 6: Two-Hit Models (CAN+LPS and CAN+PA)

```{r figure6}
# =============================================================================
# Figure 6: Two-Hit Models
# CAN-LPS (Exps 3, 5, 6 — WT, Day 4)
# CAN-PA (Exp 28 — Day 3)
# =============================================================================

# ---- CAN-LPS data ----
fig6_lps <- facts %>%
  filter(experiment %in% c(3, 5, 6), genotype == "WT",
         group %in% c("CAN6Veh", "VehLPS", "CAN6LPS")) %>%
  mutate(group = factor(group, levels = c("CAN6Veh", "VehLPS", "CAN6LPS")))
fig6_lps$BAL_IgM <- as.numeric(fig6_lps$BAL_IgM)
fig6_lps_bal <- fig6_lps %>% filter(mouseID != "WTCAN6LPS_m2")  # drowned mouse

# ---- CAN-PA data ----
fig6_pa <- facts %>%
  filter(experiment == 28) %>%
  filter(mouseID != "CA_PA_m4") %>%
  mutate(group = factor(group, levels = c("Veh_Veh", "Veh_PA", "CA_PA")))
fig6_pa$BAL_IgM <- as.numeric(fig6_pa$BAL_IgM)
fig6_pa$BAL_totalProtein <- as.numeric(fig6_pa$BAL_totalProtein)

# ---- Colors and comparisons ----
lps_colors <- c("CAN6Veh" = "gray50", "VehLPS" = "steelblue3", "CAN6LPS" = "firebrick")
pa_colors  <- c("Veh_Veh" = "gray50", "Veh_PA" = "steelblue",  "CA_PA"  = "firebrick3")
lps_comparisons <- list(c("CAN6LPS", "CAN6Veh"), c("CAN6LPS", "VehLPS"))
pa_comparisons  <- list(c("CA_PA", "Veh_Veh"), c("CA_PA", "Veh_PA"))

theme_fig6 <- function() {
  theme_pubr(base_size = 10) %+replace%
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
          plot.margin = margin(4, 4, 4, 4))
}

fig6_boxplot <- function(data, yvar, ylab, title = NULL, colors, comparisons,
                         ylimits = NULL, ybreaks = waiver()) {
  p <- ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_fig6()
  if (!is.null(ylimits)) p <- p + scale_y_continuous(limits = ylimits, breaks = ybreaks)
  p
}

# ---- Weight trajectories ----
fig6_lps_wt <- fig6_lps %>%
  dplyr::select(mouseID, group, d0weightpercent:d4weightpercent) %>%
  pivot_longer(cols = d0weightpercent:d4weightpercent,
               names_to = "Day", values_to = "WeightPct") %>%
  mutate(Day = factor(Day, levels = paste0("d", 0:4, "weightpercent"),
                      labels = paste("Day", 0:4)))
fig6_lps_wt_sum <- fig6_lps_wt %>%
  group_by(group, Day) %>%
  summarise(median = median(WeightPct, na.rm = TRUE),
            q25 = quantile(WeightPct, 0.25, na.rm = TRUE),
            q75 = quantile(WeightPct, 0.75, na.rm = TRUE), .groups = "drop")

fig6_wtA <- ggplot(fig6_lps_wt_sum,
    aes(x = Day, y = median, group = group, color = group)) +
  geom_line(linewidth = 0.8) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.15, linewidth = 0.5) +
  scale_color_manual(values = lps_colors) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Body Weight (% baseline)", title = "Weight Change: CAN + LPS", color = NULL) +
  theme_pubr(base_size = 10) +
  theme(legend.position = "top", legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.3))

fig6_pa_wt <- fig6_pa %>%
  dplyr::select(mouseID, group, d0weightpercent:d3weightpercent) %>%
  pivot_longer(cols = d0weightpercent:d3weightpercent,
               names_to = "Day", values_to = "WeightPct") %>%
  mutate(Day = factor(Day, levels = paste0("d", 0:3, "weightpercent"),
                      labels = paste("Day", 0:3)))
fig6_pa_wt_sum <- fig6_pa_wt %>%
  group_by(group, Day) %>%
  summarise(median = median(WeightPct, na.rm = TRUE),
            q25 = quantile(WeightPct, 0.25, na.rm = TRUE),
            q75 = quantile(WeightPct, 0.75, na.rm = TRUE), .groups = "drop")

fig6_wtB <- ggplot(fig6_pa_wt_sum,
    aes(x = Day, y = median, group = group, color = group)) +
  geom_line(linewidth = 0.8) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.15, linewidth = 0.5) +
  scale_color_manual(values = pa_colors) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Body Weight (% baseline)", title = "Weight Change: CAN + PA", color = NULL) +
  theme_pubr(base_size = 10) +
  theme(legend.position = "top", legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.3))

# ---- CAN-LPS endpoint panels ----
lps_lungwt <- fig6_boxplot(fig6_lps, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight",
  colors = lps_colors, comparisons = lps_comparisons, ylimits = c(0, 12.5))
lps_cells <- fig6_boxplot(fig6_lps_bal, "BALcellcount",
  ylab = expression("BAL Cells (" * 10^4 * ")"), title = "BAL Cells",
  colors = lps_colors, comparisons = lps_comparisons)
lps_pmn <- fig6_boxplot(fig6_lps_bal, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils",
  colors = lps_colors, comparisons = lps_comparisons)
lps_prot <- fig6_boxplot(fig6_lps_bal, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein",
  colors = lps_colors, comparisons = lps_comparisons)
lps_igm <- fig6_boxplot(fig6_lps_bal, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM",
  colors = lps_colors, comparisons = lps_comparisons)

# ---- CAN-PA endpoint panels ----
pa_lungwt <- fig6_boxplot(fig6_pa, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight",
  colors = pa_colors, comparisons = pa_comparisons, ylimits = c(0, 12))
pa_cells <- fig6_boxplot(fig6_pa, "BALcellcount",
  ylab = expression("BAL Cells (" * 10^4 * ")"), title = "BAL Cells",
  colors = pa_colors, comparisons = pa_comparisons)
pa_pmn <- fig6_boxplot(fig6_pa, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils",
  colors = pa_colors, comparisons = pa_comparisons)
pa_prot <- fig6_boxplot(fig6_pa, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein",
  colors = pa_colors, comparisons = pa_comparisons)
pa_igm <- fig6_boxplot(fig6_pa, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM",
  colors = pa_colors, comparisons = pa_comparisons)

# ---- Placeholders and titles ----
fig6_schematic <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "(schematic)", size = 5, color = "grey60")
title_lps <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "C. albicans + LPS",
           size = 4.5, fontface = "bold.italic")
title_pa <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "C. albicans + P. aeruginosa",
           size = 4.5, fontface = "bold.italic")

# ---- Assembly ----
row1 <- fig6_schematic + fig6_wtA + fig6_wtB + plot_layout(nrow = 1, widths = c(2, 1, 1))
row2 <- lps_lungwt + lps_cells + lps_pmn + lps_prot + lps_igm + plot_layout(nrow = 1)
row3 <- pa_lungwt + pa_cells + pa_pmn + pa_prot + pa_igm + plot_layout(nrow = 1)

fig6 <- row1 / wrap_elements(title_lps) / row2 / wrap_elements(title_pa) / row3 +
  plot_layout(heights = c(1.2, 0.08, 1, 0.08, 1)) +
  plot_annotation(tag_levels = list(
    c("A", "B", "C", "", "D", "E", "F", "G", "H", "", "I", "J", "K", "L", "M")),
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))

ggsave("Figure6_TwoHit.pdf", fig6, width = 16, height = 12)
```

## Supplemental: Two-Hit CFUs and Histology

```{r supp_twohit_cfus_histology}
# =============================================================================
# Supplemental: CFU figures for both models + CAN-LPS histology
# =============================================================================

# CAN-LPS Candida CFUs
fig6_lps_bal <- fig6_lps_bal %>% mutate(log10_lungCFU = log10(LeftLung_CanCFUs + 1))
figS_lps_cfu <- fig6_boxplot(fig6_lps_bal, "log10_lungCFU",
  ylab = expression(log[10] ~ "Lung" ~ italic("C. albicans") ~ "CFUs/g"),
  title = "CAN+LPS: Lung Candida CFUs",
  colors = lps_colors, comparisons = lps_comparisons)

# CAN-PA CFUs
fig6_pa <- fig6_pa %>%
  mutate(log10_lungCFU = log10(LeftLung_CanCFUs + 1),
         log10_psaCFU  = log10(Lung_PSA_CFU + 1))
figS_pa_canCfu <- fig6_boxplot(fig6_pa, "log10_lungCFU",
  ylab = expression(log[10] ~ "Lung" ~ italic("C. albicans") ~ "CFUs/g"),
  title = "CAN+PA: Lung Candida CFUs",
  colors = pa_colors, comparisons = pa_comparisons)
figS_pa_psaCfu <- fig6_boxplot(fig6_pa, "log10_psaCFU",
  ylab = expression(log[10] ~ "Lung" ~ italic("P. aeruginosa") ~ "CFUs/g"),
  title = "CAN+PA: Lung Pseudomonas CFUs",
  colors = pa_colors, comparisons = pa_comparisons)

figS_cfus <- figS_lps_cfu + figS_pa_canCfu + figS_pa_psaCfu +
  plot_layout(nrow = 1) +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))
ggsave("FigS_TwoHit_CFUs.pdf", figS_cfus, width = 10, height = 4)

# CAN-LPS Histology
histology2hit <- histology2hit %>%
  filter(Reviewer == "Lokesh",
         group %in% c("CAN6Veh", "VehLPS", "CAN6LPS")) %>%
  mutate(group = factor(group, levels = c("CAN6Veh", "VehLPS", "CAN6LPS")))

figS_histo_2hit <- ggplot(histology2hit,
    aes(x = group, y = ScorePerField, color = group)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
  geom_jitter(width = 0.15, size = 1.2, alpha = 0.6) +
  stat_compare_means(comparisons = lps_comparisons, size = 3,
                     label = "p.format", bracket.size = 0.4) +
  scale_color_manual(values = lps_colors) +
  labs(x = NULL, y = "Histology Score", title = "CAN+LPS Histology Score") +
  theme_fig6()

figS_histo_images <- ggplot() + theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "(histology images)", size = 5, color = "grey60")

figS_histo <- figS_histo_2hit + figS_histo_images +
  plot_layout(widths = c(0.6, 1)) +
  plot_annotation(tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 14, face = "bold")))
ggsave("FigS_TwoHit_Histology.pdf", figS_histo, width = 8, height = 4)
```

---

## Supplemental: Yeast-Lock + *P. aeruginosa* (Exp 39)

```{r supp_yeastlock_pa}
# =============================================================================
# Supplemental: Yeast-Lock vs WT C. albicans + P. aeruginosa (Exp 39)
# =============================================================================

fig_ylpa <- facts %>%
  filter(experiment == 39, genotype == "WT") %>%
  mutate(group = case_when(
    group == "TT21_PA"  ~ "Wild-type + PA",
    group == "TNRG1_PA" ~ "Yeast-lock + PA",
    TRUE ~ group)) %>%
  mutate(group = factor(group, levels = c("Wild-type + PA", "Yeast-lock + PA")))
fig_ylpa$BAL_IgM <- as.numeric(fig_ylpa$BAL_IgM)
fig_ylpa <- fig_ylpa %>%
  mutate(log10_lungCFU = log10(LeftLung_CanCFUs + 1),
         log10_psaCFU  = log10(Lung_PSA_CFU + 1))

ylpa_colors <- c("Wild-type + PA" = "forestgreen", "Yeast-lock + PA" = "goldenrod3")
ylpa_comparisons <- list(c("Wild-type + PA", "Yeast-lock + PA"))

theme_ylpa <- function() {
  theme_pubr(base_size = 10) %+replace%
    theme(legend.position = "none",
          axis.title.y = element_text(size = 9, angle = 90),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
          plot.margin = margin(4, 4, 4, 4))
}

ylpa_boxplot <- function(data, yvar, ylab, title = NULL,
                         ylimits = NULL, ybreaks = waiver()) {
  p <- ggplot(data, aes(x = group, y = .data[[yvar]], color = group)) +
    geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.5) +
    geom_beeswarm(size = 1.2, cex = 2) +
    stat_compare_means(comparisons = ylpa_comparisons, size = 3,
                       label = "p.format", bracket.size = 0.4) +
    scale_color_manual(values = ylpa_colors) +
    labs(x = NULL, y = ylab, title = title) +
    theme_ylpa()
  if (!is.null(ylimits)) p <- p + scale_y_continuous(limits = ylimits, breaks = ybreaks)
  p
}

# Weight trajectory
ylpa_wt_long <- fig_ylpa %>%
  dplyr::select(mouseID, group, d0weightpercent:d3weightpercent) %>%
  pivot_longer(cols = d0weightpercent:d3weightpercent,
               names_to = "Day", values_to = "WeightPct") %>%
  mutate(Day = factor(Day, levels = paste0("d", 0:3, "weightpercent"),
                      labels = paste("Day", 0:3)))
ylpa_wt_sum <- ylpa_wt_long %>%
  group_by(group, Day) %>%
  summarise(median = median(WeightPct, na.rm = TRUE),
            q25 = quantile(WeightPct, 0.25, na.rm = TRUE),
            q75 = quantile(WeightPct, 0.75, na.rm = TRUE), .groups = "drop")

ylpa_A <- ggplot(ylpa_wt_sum,
    aes(x = Day, y = median, group = group, color = group)) +
  geom_line(linewidth = 0.8) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.15, linewidth = 0.5) +
  scale_color_manual(values = ylpa_colors) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Body Weight (% baseline)", title = "Weight Change", color = NULL) +
  theme_pubr(base_size = 10) +
  theme(legend.position = "top", legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.3))

ylpa_B <- ylpa_boxplot(fig_ylpa, "LungWeightNormalized",
  ylab = "Lung Weight / BW (mg/g)", title = "Normalized Lung Weight")
ylpa_C <- ylpa_boxplot(fig_ylpa, "BALcellcount",
  ylab = expression("BAL Cells (" * 10^4 * ")"), title = "BAL Cells")
ylpa_D <- ylpa_boxplot(fig_ylpa, "BAL_Neutrophils_total",
  ylab = expression("BAL PMNs (" * 10^4 * ")"), title = "BAL Neutrophils")
ylpa_E <- ylpa_boxplot(fig_ylpa, "log10_lungCFU",
  ylab = expression(log[10] ~ italic("C. albicans") ~ "CFUs/g"), title = "Lung Candida CFUs")
ylpa_F <- ylpa_boxplot(fig_ylpa, "log10_psaCFU",
  ylab = expression(log[10] ~ italic("P. aeruginosa") ~ "CFUs/g"), title = "Lung Pseudomonas CFUs")
ylpa_G <- ylpa_boxplot(fig_ylpa, "BAL_totalProtein",
  ylab = "BAL Protein (mg/ml)", title = "BAL Protein")
ylpa_H <- ylpa_boxplot(fig_ylpa, "BAL_IgM",
  ylab = "BAL IgM (ng/ml)", title = "BAL IgM")

row1 <- ylpa_A + ylpa_B + ylpa_C + ylpa_D + plot_layout(nrow = 1)
row2 <- ylpa_E + ylpa_F + ylpa_G + ylpa_H + plot_layout(nrow = 1)

figS_ylpa <- row1 / row2 +
  plot_annotation(
    title = expression(bold("Yeast-Lock vs Wild-Type" ~ italic("C. albicans") ~ "+ " ~ italic("P. aeruginosa"))),
    tag_levels = "A",
    theme = theme(plot.title = element_text(size = 12, hjust = 0.5),
                  plot.tag = element_text(size = 14, face = "bold")))

ggsave("FigS_YeastLock_PA.pdf", figS_ylpa, width = 14, height = 7)
```
